### 2016-06-14 cvclust.tex
revise through sec 4.1

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{float}
%\usepackage{footnote}
%\usepackage[bottom]{footmisc}
\usepackage[singlelinecheck=false]{caption}

\usepackage{graphicx,psfrag,epsf}
\usepackage[space]{grffile}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{1}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{-.3in}%
\addtolength{\topmargin}{-.8in}%

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{claim}{Claim}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}

% operators
\DeclareMathOperator*{\argmin}{arg\,min}
\newcommand{\CV}{\operatorname{CV}}
\newcommand{\PE}{\operatorname{PE}}
\newcommand{\E}{\operatorname{E}}
\DeclareMathOperator*{\cov}{cov}
\DeclareMathOperator*{\tr}{tr}
\DeclareMathOperator*{\var}{var}

% convergence
\newcommand{\toas}{\overset{\mathit{a.s.}}{\to}}

\newcommand{\OhP}{O_p}

% sets
\newcommand{\R}{\mathbb{R}}
\newcommand{\sA}{\mathcal{A}}
\newcommand{\sbA}{\mathcal{\bar A}}
\newcommand{\sC}{\mathcal{C}}
\newcommand{\sE}{\mathcal{E}}
\newcommand{\sN}{\mathcal{N}}
\newcommand{\sX}{\mathcal{X}}

% scalars
\newcommand{\bpi}{\bar \pi}

% vectors
\newcommand{\muX}{\mu^{X}}
\newcommand{\muY}{\mu^{Y}}
\newcommand{\bmuX}{\bar \mu^{X}}
\newcommand{\bmuY}{\bar \mu^{Y}}
\newcommand{\hmuY}{\hat \mu^{Y}}

% matrices
\newcommand{\dataX}{\mathfrak{X}}
\newcommand{\SigmaY}{\Sigma^Y}
\newcommand{\Xtrain}{X_{\text{train}}}
\newcommand{\Ytrain}{Y_{\text{train}}}
\newcommand{\Xtest}{X_{\text{test}}}
\newcommand{\Ytest}{Y_{\text{test}}}

% class labels
\newcommand{\hGX}{\hat G^{X}}
\newcommand{\hGY}{\hat G^{Y}}

\begin{document}

%\bibliographystyle{natbib}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if1\blind
{
  \title{\bf Estimating the number of clusters using cross-validation}
  \author{Wei Fu and Patrick O. Perry \\
  Stern School of Business, New York University}
  \maketitle
} \fi

\if0\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Estimating the number of clusters using cross-validation}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
Many clustering methods, including $k$-means, require the user to specify the
number of clusters as an input parameter. A variety of methods have been
devised to choose the number of clusters automatically, but they often rely on
strong modeling assumptions. We propose a data-driven approach to estimate the
number of clusters based on a novel form of cross-validation. This differs
from ordinary cross-validation, because clustering is fundamentally an
unsupervised learning problem. Simulation and real data analysis results show
that our proposed method outperforms existing methods, especially in
high-dimensional settings with heavy-tailed data.
\end{abstract}

\noindent%
{\it Keywords:} clustering, unsupervised learning, model selection
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!
\section{Introduction}
\label{sec:intro}

A clustering procedure segments a collection of items into smaller groups,
with the property that items in the same group are more similar to each other
than items in different groups \citep{hartigan1975clustering}.  Such
procedures are used in two main applicaitons: (a) exploratory analysis, where clusters
reveal homogeneous sub-groups within a large sample; (b) data
reduction, where high-dimensional item attribute vectors get reduced discrete
cluster labels \citep{jain1999data}.

With many clustering methods, including the popular $k$-means clustering
procedure, the user must specify $k$, the number of clusters
\citep{jain2010data}.  One popular ad-hoc device for selecting the number of
clusters is to use an analogue of the principal components scree plot: plot
the within-cluster dispersion $W_k$, as a function of the number of clusters
$k$, looking for an ``elbow'' in the plot.  This approach is simple and often
performs well, but it requires subjective judgment as to where the elbow is
located, and as we demonstrate in Appendix~\ref{sec:elbow-fail}, the approach
can easily fail.  In this report, we propose a new method to choose the number
of clusters automatically.

The problem of choosing $k$ has been well-studied, and dozens of methods have
been proposed~\citep{chiang2010intelligent,fujita2014non}. The main difficulty
in choosing $k$ is that clustering is fundamentally an ``unsupervised''
learning problem, meaning that there is no obvious way to use ``prediction
ability'' to drive the model selection \citep{hastie2009elements}.  Most
existing methods for choosing $k$ instead rely on on explicit or implicit
assumptions about the data distribution, including it shape, scale, and
correlation structure.

Several authors advocate choosing $k$ by performing a sequence of hypothesis
tests will null and alternative hypotheses of the form $H_0 : k = k_0$ and
$H_1: k > k_0$, starting with $k_0 = 1$ and proceeding sequentially with
higher values of $k_0$ until a test fails to reject $H_0$. The gap statistic
method typifies this class of methods, with a test statistic that measures the
within-cluster dispersion relative to what is expected under a reference
distribution \citep{tibshirani2001estimating}.
%Similar methods include those developed by \citet{TODO}.

Other authors have proposed choosing $k$ by using information
criteria.  For example, \citet{sugar2003finding} proposed an approach that
minimizes the estimated ``distortion'', the average distance per dimension.
Likewise, \citet{fraley2002model}'s model-based method fits Gaussian mixture model
models to the data, then selects the number of mixture components, $k$, using
the Bayesian Information Criterion (BIC).

A third set of approaches is based on the idea of ``stability'', that clusters
are meaningful if they manifest in multiple independent samples from the same
population. \citet{ben2001stability}, \citet{tibshirani2005cluster},
\citet{wang2010consistent} and \citet{fang2012selection} developed methods
based on this idea.

The procedure we propose in this report is based on a form of
cross-validation, and it is adaptive to the characteristics of the data
distribution. The essential idea is to devise a way to measure a form of
internal prediction error associated each choose of $k$, and then choose the
$k$ with the smallest associated error. We describe this method in detail in
Section~\ref{sec:meth}.  In Section~\ref{sec:self-consistent}, we prove that
our method is self-consistent. Then, in Section~\ref{sec:theory-gaussian}, we
analyze the performance of our method in the presence of Gaussian noise. The
theoretical analysis shows that the performance of our method degrades in the
presence of correlated noise; to fix this, we propose a correction for
correlated noise in Section~\ref{sec:corr-correct}. In
Sections~\ref{sec:simulations} and~\ref{sec:empirical-validation}, we
demonstrate that our method is competitive with other state-of-the-art
procedures in both simulated and real data sets. Then, in
Section~\ref{sec:application-yeast}, we apply our method to a Yeast cell cycle
dataset. We conclude with a short discussion in Section~\ref{sec:discussion}.

\section{Cross-validation for clustering}
\label{sec:meth}

\subsection{Problem statement}

Suppose that we are given a data matrix with $N$ rows and $P$ columns, and we
are tasked with choosing an appropriate number $k$ of clusters to use for
performing $k$-means clustering on the rows of the data matrix. Recall that
the $k$-means procedure takes a set of observations $\{ x_1, \dotsc ,x_n \}$
and finds a set of $k$ or cluster centers $A = \{ a_1, \dotsc, a_k \} $
minimizing the within cluster dispersion
\[
  W(A) = \sum_{i=1}^{n} \min_{a \in A} \|x_i - a\|^2.
\]
This implicitly defines a cluster assignment rule
\[
  g(x) = \argmin_{g \in \{1, \dotsc, k\}} \|x - a_g\|^2,
\]
with ties broken arbitrarily.
%We will assume that the $k$-means procedure
%finds an optimal choice of , $A$, but we will not assume that this solution is
%unique.

We can consider the problem of choosing $k$, the number of clusters, to be a
model selection problem. In other domains, especially supervised learning
problems like regression and classification, cross-validation is popular for
performing model selection.
In these settings, the data comes in the form of $N$
predictor-response pairs, $(X_1, Y_1), \dotsc, (X_N, Y_N)$, with $X_i \in
\R^{p}$ and $Y_i \in \R^{q}$.  The data can be represented as a matrix with
$N$ rows and $p + q$ columns.  We partition the data into $K$ hold-out
``test'' subsets, with $K$ typically chosen to be $5$ or $10$.  For each
``fold'' $r$ in the range $1, \dotsc, K$, we permute the rows of the data
matrix to get $\dataX$, a matrix with the $r$th test subset as its trailing
rows.  We partition $\dataX$ as
\[
  \dataX =
  \begin{bmatrix}
    \Xtrain & \Ytrain \\
    \Xtest  & \Ytest
  \end{bmatrix}.
\]
We use the training rows $[ \Xtrain\ \Ytrain ]$ to fit a regression model
$\hat Y = \hat Y(X)$, and then evaluate the performance of this model on the
test set, computing the cross-validation error $\|\Ytest - \hat Y(\Xtest)\|^2$
or some variant thereof.  We choose the model with the smallest
cross-validation error, averaged over all $K$ folds.

In unsupervised learning problems like factor analysis and clustering, the
features of the observations are not naturally partitioned into ``predictors''
and ``responses'', so we cannot directly apply the cross-validation procedure
described above.  For factor analysis, there are at least two versions of
cross-validation.  \citet{wold78cross} proposed a ``speckled'' holdout, where
in each fold we leave out a subset of the elements of the data matrix.  Wold's
procedure works well empirically, but does not have any theoretical support,
and it requires a factor analysis procedure that can handle missing data.
\citet{owen2009bi} proposed a scheme called ``bi-cross-validation'' wherein
each fold designates a subset of the data matrix columns to be response and a
subset of the rows to be test data.  This generalized a procedure due to
\citet{gabriel2002biblot}, who proposed holding out a single column and a
single row at each fold.

%% [POP] This paragraph doesn't fit into the section.  Maybe put it in the introduction?
%% 
%% Cross-Validation technique, one of the most commonly used and popular model
%% selecting method in supervised learning, can not be used naively in
%% unsupervised learning context. Since prediction error of new observation is
%% calculated by its distance to the nearest cluster center, more cluster centers
%% means much tighter fit to the feature space, and hence smaller distance
%% (prediction error) of observation to its nearest cluster center. This holds
%% even when the prediction is evaluated on an independent test set
%% \citep{hastie2009elements}. Therefore, CV prefers larger $k$ if we do the CV
%% naively.

In the sequel, we extend Gabriel cross-validation to the problem of selecting
the number of clusters, $k$, automatically, and we provide theoretical and
empirical support analogous to the consistency results proved by
\citet{owen2009bi}.

\subsection{Gabriel cross-validation}
\label{sec:gabriel-cv-algorithm}

Our version of Gabriel cross validation for clustering works by performing a
sequence of ``folds'' over the data. We use these folds to estimate a version
of prediction error (cross-validation error) for each possible value of $k$;
we then choose the value $\hat k$ with the smallest cross-validation error.

In each fold of our cross-validation procedure, we permute the rows and
columns of the data matrix and then partition the rows and columns as $N = n +
m$ and $P = p + q$ for positive integers $n$, $m$, $p$, and $q$.  We treat
the first $p$ columns as ``predictors'' and the last $q$ columns as
``responses''; similarly, we treat the first $n$ rows as ``training'' and the
last $m$ rows as ``test''.  In block form, the permuted data matrix is
\[
  \dataX
  =
  \begin{bmatrix}
    \Xtrain & \Ytrain \\
    \Xtest  & \Ytest
  \end{bmatrix},
\]
where
$\Xtrain \in \R^{n \times p}$,
$\Ytrain \in \R^{n \times q}$,
$\Xtest \in  \R^{m \times p}$,
and
$\Ytest \in  \R^{m \times q}$.

Given such a partition of $\dataX$, we perform four steps for each value of
$k$, the number of clusters:
\begin{enumerate}
  \item \label{step:gabriel-cluster}
    \textbf{Cluster:}
    Cluster $Y_{1}, \dotsc, Y_n$, the rows of $\Ytrain$, yielding the
    assignment rule $\hGY : \R^q \to \{ 1, \dotsc, k \}$ and the
    cluster means $\bmuY_1, \dotsc, \bmuY_k$.  Set $\hGY_i = \hGY(Y_i)$ to
    be the assigned cluster for row $i$.
  \item \label{step:gabriel-classify}
    \textbf{Classify:}
    Take $X_{1}, \dotsc, X_n$, the rows of $\Xtrain$ to be predictors,
    and take $\hGY_1, \dotsc, \hGY_n$ to be corresponding class labels.  Use
    the pairs $\{ (X_i, \hGY_i) \}_{i=1}^{n}$ to train a classifier
    $\hGX : \R^p \to \{ 1, \dotsc, k \}$.
  \item \label{step:gabriel-predict}
    \textbf{Predict:}
    Apply the classifier to $X_{n+1}, \dotsc, X_{n+m}$, the rows of
    $\Xtest$, yielding predicted classes $\hGX_i = \hGX(X_i)$ for
    $i = n+1, \dotsc, n+m$.  For each value of $i$ in this range, compute
    predicted response $\hat Y_i = \bmuY(\hGX_i)$, where
    $\bmuY(g) = \bmuY_g$.
  \item \label{step:gabriel-evaluate}
    \textbf{Evaluate:}
    Compute the cross-validation error
    \[
      \CV(k) = \frac{1}{m} \sum_{i=n+1}^{n+m} \|Y_i - \hat Y_i\|^2,
    \]
    where $Y_{n+1}, \dotsc, Y_{n+m}$ are the rows of $\Ytest$.
\end{enumerate}
\noindent
In principle, we could use any clustering and classification methods in
steps~\ref{step:gabriel-cluster} and~\ref{step:gabriel-classify}.  In this
report, we use $k$-means as the clustering algorithm.  For the classification
step, we compute the mean value of $X$ for each class; we assign an
observation to class $g$ if that class has the closest mean (randomly breaking
ties between classes).  The classification step is equivalent to linear
discriminant analysis with equal class priors and identity noise covariance
matrix.

To choose the folds, we randomly partition the rows and columns into $K$ and
$L$ subsets, respectively.  Each fold is indexed by a pair $(r,s)$ of
integers, with $r \in \{1, \dotsc, K\}$ and $s \in \{1, \dotsc, L\}$.  Fold
$(r,s)$ treats the $r$th row subset as ``test'', and the $s$th column subset
as ``response''.  We typically take $K = 5$ and $L = 2$.  For the number of
clusters, we select the value of $k$ that minimizes the average of $\CV(k)$
over all $K \times L$ folds (choosing the smallest value of $k$ in the event
of a tie).

TODO: describe theoretical support

\section{Self-consistency}
\label{sec:self-consistent}

An important property of any estimation procedure is that in the absence of of
noise, the procedure correctly estimates the truth. This property is called
``self-consistency'' \citep{tarpey96}. We will now show that Gabriel
cross-validation is self-consistent. That is, in the absence of noise, the
Gabriel cross-validation procedure finds the optimal number of clusters.

It will suffice to prove self-consistency for a single fold of the
cross-validation procedure.  As in in section~\ref{sec:gabriel-cv-algorithm}
we assume that the $P$ variables of the data set have been partitioned into
$p$ predictor variables represented in vector~$X$ and $q$ response variables
represented in vector~$Y$.  The $N$ observations have been divided into two
sets: $n$ train observations and $m$ test observations.  The following theorem
gives conditions for Gabriel CV to recover the true number of clusters in the
absence of noise.

\begin{theorem}\label{thm:self-consistency}

Let $\{ (X_i, Y_i) \}_{i=1}^{n+m}$ be the data from a single fold of Gabriel
cross-validation.  For any $k$, let $CV(k)$ be the cross-validation error for
this fold, computed as described in Section~\ref{sec:gabriel-cv-algorithm}.
We will assume that there are $K$ true centers $\mu(1), \dotsc,\mu(K)$, with
the $g$th cluster center partitioned as $\mu(g) = \bigl(\muX(g),
\muY(g)\bigr)$ for $g = 1, \dotsc, K$.  Suppose that
\begin{enumerate}[label=(\roman*)]
  \item \label{asn:self-consistency-noiseless}
    Each observation $i$ has a true cluster $G_i \in \{ 1, \dotsc, K \}$.
    There is no noise, so that $X_i = \muX({G_i})$ and $Y_i = \muY(G_i)$ for
    $i = 1, \dotsc, n+m$.
  \item \label{asn:self-consistency-distinct-mux}
    The vectors $\muX(1), \dotsc,\muX(K)$ are all distinct.
  \item \label{asn:self-consistency-distinct-muy}
    The vectors $\muY(1), \dotsc,\muY(K)$ are all distinct.
  \item \label{asn:self-consistency-train}
    The training set contains at least one member of each cluster: for all $g$
    in the range $1, \dotsc, K$, there exists at least one $i$ in the range
    $1, \dotsc, n$ such that $G_i = g$.
  \item \label{asn:self-consistency-test}
    The test set contains at least one member of each cluster: for all $g$ in
    the range $1, \dotsc, K$, there exists at least one $i$ in the range $n+1,
    \dotsc, n+m$ such that $G_i = g$.
\end{enumerate}
Then $CV(k) < CV(K)$ for $k < K$, and $CV(k) = CV(K)$ for $k > K$, so that
Gabriel CV correctly chooses $k = K$.
\end{theorem}

The result statement of Theorem~\ref{thm:self-consistency} follows from 
Lemmas~\ref{lem:self-consistency1} and~\ref{lem:self-consistency2}, which we
will now state and prove.

\begin{lemma}\label{lem:self-consistency1}
Suppose that the assumptions of Theorem~\ref{thm:self-consistency} are in
force.  If $k < K$, then $\CV(k) > 0$.
\end{lemma}
\begin{proof}
By definition,
\[
  \CV(k)
    =
      \sum_{i=n+1}^{n+m}
        \| Y_i - \bmuY (\hGX_i) \|^2,
\]
where $\bmuY(g)$ is the center of cluster $g$ returned from applying $k$-means
to $Y_1, \dotsc, Y_n$.  Assumptions~\ref{asn:self-consistency-noiseless}
and~\ref{asn:self-consistency-test}, imply that as $i$ ranges over the test
set $n+1, \dotsc, n+m$, the response $Y_i$ ranges over all distinct values in
$\{ \muY(1), \dotsc, \muY(K) \}$.
Assumption~\ref{asn:self-consistency-distinct-muy} implies that there are
exactly $K$ such distinct values.  However, there are only $k$ distinct values
of $\bmuY(g)$.  Thus, at least one summand
\(
  \| Y_i - \bmuY(\hGX_i) \|^2
\)
is nonzero.  Therefore,
\(
  \CV(k) > 0.
\)
\end{proof}

\begin{lemma}\label{lem:self-consistency2}
Suppose that the assumptions of Theorem~\ref{thm:self-consistency} are in
force.  If $k \geq K$, then $\CV(k) = 0$.
\end{lemma}
\begin{proof}
From assumptions~\ref{asn:self-consistency-noiseless},
\ref{asn:self-consistency-distinct-muy},
and~\ref{asn:self-consistency-train}, we know the cluster centers
gotten from applying $k$-means to $Y_1, \dotsc, Y_n$ must include
$\muY(1), \dotsc, \muY(K)$.  Without loss of generality, suppose that
$\bmuY(g) = \muY(g)$ for $g = 1, \dotsc, K$.  This implies that
$\hGY_i = G_i$ for $i = 1, \dotsc, n$.  Thus, employing
assumption~\ref{asn:self-consistency-noiseless} again, we get that
$\bmuX(g) = \muX(g)$ for $g = 1, \dotsc, K$.

Since assumption~\ref{asn:self-consistency-distinct-mux} ensures that
$\muX(1), \dotsc, \muX(K)$ are all distinct, we must have that $\hGX_i = G_i$
for all $i = 1, \dotsc, m+n$.  In particular, this implies that $\bmuY(\hGX_i)
= Y_i$ for $i = 1, \dotsc, m+n$, so that $\CV(k) = 0$.
\end{proof}

\section{Analysis under Gaussian noise}
\label{sec:theory-gaussian}

\subsection{Single cluster in two dimensions}

Theorem~\ref{thm:self-consistency} tells us that the Gabriel cross-validation
method recovers the true number of clusters when the noise is negligible.
While this result gives us some assurance that the procedure is well-behaved,
we can bolster our confidence and gain insight into its workings by analyzing
its behavior in the presence of noise.  We first study the case of a single
cluster in two dimensions with correlated Gaussian noise.

\begin{theorem}
Suppose that $\{ (X_i, Y_i) \}_{i=1}^{n + m}$ is data from a single fold
of Gabriel cross-validation, where each $(X,Y)$ pair in $\R^2$ is an
independent draw from a mean-zero multivariate normal distribution with unit
marginal variances and correlation $\rho$.  In this case, the data are drawn
from a single cluster; the true number of clusters is~$1$.  If $|\rho| < 1/2$,
and $k > 1$, then $\CV(1) < \CV(k)$ with probability tending to one as $m$
and $n$ increase.
\end{theorem}

\begin{proof}

Throughout the proof we will assume that $\rho \geq 0$; a similar argument
holds with minor modification when $\rho < 0$.

Set $\hGY_1, \dotsc, \hGY_n$ to be the cluster labels gotten from applying
$k$-means to $Y_1, \dotsc, Y_n$. Denote the cluster means by $\bmuY_1 \leq
\bmuY_2 \leq \dotsb \leq \bmuY_k$.  Pollard's \citeyearpar{pollard1981strong}
strong consistency theorem for $k$-means implies that for large $n$, the
cluster centers are close to population clusters centers $a_1 < a_2 < \dotsb <
a_k$. Specifically, $\bmuY_j = a_j + \OhP(n^{-1/2})$.  Since the distribution
of $Y$ is symmetric, the population centers $a_1, a_2, \dotsc, a_k$ are
symmetric about the origin.

For $j$ in $1, \dotsc, k$, set
\[
  \bmuX_j = \frac{\sum_{i=1}^{n} 1\{ \hGY_i = j \} X_i}
                 {\sum_{i=1}^{n} 1\{ \hGY_i = j \}}.
\]
The classification rule $\hGX$ is defined by
\(
  \hGX(X) = \argmin_j \| \bmuX_j - X \|.
\)
Denote the boundaries between the population clusters as
$b_j = (a_j + a_{j+1})/2$ for $j = 1, \dotsc, k-1$.
Set $b_0 = -\infty$ and $b_k = +\infty$. 
Then, $\bmuX_j$ is within $\OhP(n^{-1/2})$ of the following expectation:
\begin{align*}
  \E(X \mid b_j \leq Y \leq b_{j+1})
    &= \E\{ E(X \mid Y) \mid b_j \leq Y \leq b_{j+1}\} \\
    &= \E\{ \rho Y \mid b_j \leq Y \leq b_{j+1}\} \\
    &= \rho \E\{ Y \mid b_j \leq Y \leq b_{j+1}\} \\
    &= \rho a_j.
\end{align*}
That is, $\bmuX_j = \rho a_j + \OhP(n^{-1/2})$. For $j = 1, \dotsc, k-1$,
the boundary between sample the classification based on $X$ to
labels $j$ and $j+1$ is
$(\bmuX_j + \bmuX_{j+1}) /2 = \rho b_j + \OhP(n^{-1/2})$.

Set $\hGX_i = \hGX(X_i)$.
The cross-validation error is
\begin{align*}
  \CV(k)
  &= \frac{1}{m} \sum_{i=n+1}^{n+m} \|Y_i - \hat Y_i\|^2
  \\
  &= \frac{1}{m} \sum_{i=n+1}^{n+m}
  \sum_{j=1}^{k} \|Y_i - \bmuY_j\|^2 1\{ \hGX_i = j \}
  \\
  &=
  \sum_{j=1}^{k}
    \pi_j
    \E[ \|Y - a_j\|^2 \mid \rho b_j < X < \rho b_{j+1}]
  + \OhP(n^{-1/2})
  + \OhP(m^{-1/2}),
\end{align*}
where $\pi_j = \Pr(\rho b_j < X < \rho b_{j+1}).$

For $j = 1, \dotsc, k$,
set
\(
  \hmuY_j =
  \E[ Y  \mid \rho b_j < X < \rho b_{j+1}].
\)
Note that
\[
  \CV(1)
    = 
  \sum_{j=1}^{k}
    \pi_j
    \E[ \|Y - 0\|^2 \mid \rho b_j < X < \rho b_{j+1}]
  + \OhP(n^{-1/2})
  + \OhP(m^{-1/2}).
\]
Thus, the difference in cross-validation errors is
\begin{align*}
\CV(k) - \CV(1)
  &=
  \sum_{j=1}^{k}
    \pi_j
    a_j
    (a_j - 2 \hmuY_j)
  + \OhP(n^{-1/2})
  + \OhP(m^{-1/2}).
\end{align*}
For arbitrary $j$,
\begin{align*}
  \hmuY_j &=
  \E[ E(Y\mid X)  \mid \rho b_j < X < \rho b_{j+1}]
\\
  &= \rho \E[ X  \mid \rho b_j < X < \rho b_{j+1}].
\end{align*}
Since $0 \leq \rho \leq 1$,
in cases where $0 \leq b_j < b_{j+1}$, we have that
$\hmuY_j \leq \rho a_j$; similarly, when $b_j < b_{j+1} \leq 0$, it follows
that $\hmuY_j \geq \rho a_j$. In either of these two situations, if $\rho <
1/2$, then
\[
  a_j (a_j - 2 \hmuY_j) > 0.
\]
The last situation to consider is when $b_j < 0 < b_{j+1}$, in which case
$b_j = - b_{j+1}$; here, $\hmuY_j = a_j = 0$.
Putting this all together, we have that as $n$ and $m$ tend to infinity,
the probability that
\(
  \CV(k) > \CV(1)
\)
tends to one.
\end{proof}

We confirm this result with a simulation.  We perform $10$ replicates.  In each
replicate, we generate $20000$ observations from a mean-zero bivariate normal
distribution with unit marginal variances and correlation $\rho$.  We perform
a single $2 \times 2$ fold of Gabriel cross-validation and report the
cross-validation mean squared error for the number of clusters $k$ ranging
from $1$ to $5$.  Figure~\ref{fig:nullcorr-equal} shows the cross-validation
errors for all $10$ replicates.  The simulation demonstrates that in the
Gabriel cross-validation criterion chooses the correct answer $k = 1$ whenever
$\rho < 0.5$; the criterion chooses $k \geq 2$ clusters whenever $|\rho| > 0.5$.

\begin{figure}[H]
\centering
\includegraphics[width=3in]{demo/nullcorr/equal.pdf}
\caption{Cross-validation error on $10$ replicates, with the number of
clusters $k$ ranging from $1$ to $5$.  Data is generated from two-dimensional
multivariate normal distribution with correlation $\rho$.  The Gabriel
cross-validation criterion chooses the correct answer $k = 1$ whenever
$|\rho| < 0.5$; the criterion chooses $k \geq 2$ clusters whenever $|\rho| > 0.5$.}
\label{fig:nullcorr-equal}
\end{figure}

Intuitively, when the correlation is high, the response feature, $Y$, looks
similar to the predictor feature, $X$. Prediction error on $X$ always
decreases with larger $k$. Thus, when the correlation is high, the prediction
error for $Y$ will also decrease with larger $k$. This explains why
cross-validation breaks down in the presense of strong correlation.

\subsection{Single Cluster in higher dimensions}

\begin{proposition}
Suppose that $\{ (X_i, Y_i) \}_{i=1}^{n + m}$ is data from a single fold
of Gabriel cross-validation, where each $(X,Y)$ pair in $\R^{p+q}$ is an
independent draw from a mean-zero multivariate normal distribution with 
covariance matrix $\Sigma_{XY}= \left( \begin{smallmatrix} \Sigma_{XX} & \Sigma_{XY} \\ 
 \Sigma_{YX} & \Sigma_{YY} \end{smallmatrix}\right)$, with $\Sigma_{YY}$ has leading 
eigenvalue $\lambda_1$ and corresponding eigenvector $u_1$. In this case, the data are drawn
from a single cluster; the true number of clusters is~$1$.  If  $\frac{\sqrt{\lambda_1}}{2}
 > \frac{u^T_1\Sigma_{YX}\Sigma_{XY}u_1}{\sqrt{u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1}}$,
then $\CV(1) < \CV(2)$ with probability tending to one as $m$ and $n$ increase.
\end{proposition}

\begin{proof}
Let $X$ and $Y$ be jointly multivariate normal distributed with mean $\mathbf{0}$ and covariance matrix $\Sigma_{XY}$, i.e.
\[	(X,Y) \sim \mathcal{N} \left( \mathbf{0}, \Sigma_{XY}\right)	\]
where $\Sigma_{XY}=\begin{bmatrix} \Sigma_{XX} & \Sigma_{XY} \\  \Sigma_{YX} & \Sigma_{YY} \end{bmatrix}$.

Let $\Sigma_{YY} = U \Lambda U^T$  be the eigendecomposition of $\Sigma_{YY}$, with leading eigenvalue $\lambda_1$ and corresponding eigenvector $u_1$. Then the centroid of $k$-means applying on $(y_1,..,y_n)$ is on the first PC 
of $Y$,\[	E(u^T_1 Y|u^T_1 Y>0) = \bar{\mu}^Y_1 =\sqrt{2 \lambda_1/\pi}u_1\] and 
\[	E(u^T_1 Y|u^T_1 Y<0) = \bar{\mu}^Y_2 =-\sqrt{2 \lambda_1/\pi}u_1\]
 where $u^T_1 Y \sim \mathcal{N}(0,\lambda_1)$.

To compute $\bar{\mu}^X_1 = E(X|u^T_1 Y>0)$, we need to know the conditional distribution $X|u^T_1 Y$. Since $(X,Y)$ has multivariate normal distribution, $(X,u^T_1 Y)$ also has a multivariate normal distribution with mean $\mathbf{0}$ and covariance matrix
$$\Sigma_{X,u^T_1 Y}=\begin{bmatrix} \Sigma_{XX} & \Sigma_{XY} u_1 \\  u^T_1 \Sigma_{YX} & \lambda_1 \end{bmatrix}$$
The conditional distribution $X|u^T_1 Y$ is hence normal with mean
 $$\mu_{X|u^T_1 Y} = \Sigma_{XY} u_1 \lambda^{-1}_1 u^T_1 Y $$
Therefore, 
\begin{align}
\bar{\mu}^X_1 &= E(X \mid u^T_1 Y>0) \nonumber \\ \nonumber
 			  &= E\left(E[X \mid u^T_1 Y] \mid u^T_1 Y>0\right) \\ \nonumber
 			  &=  E\left(\Sigma_{XY} u_1 \lambda^{-1}_1 u^T_1 Y \mid u^T_1 Y>0\right)\\ \nonumber
 			  &= \lambda^{-1}_1 \Sigma_{XY}u_1 E(u^T_1 Y \mid u^T_1 Y>0) \\ \nonumber
 			  &= \lambda^{-1}_1 \Sigma_{XY}u_1 \sqrt{2 \lambda_1/\pi} \\ \nonumber
 			  &= \sqrt{2 / \lambda_1 \pi} \Sigma_{XY}u_1
\end{align}
Similar calculation yields $\bar{\mu}^X_2 = -\sqrt{2 / \lambda_1 \pi} \Sigma_{XY}u_1$.
The decision rule to classify any observed value of $X$ to $\bar{\mu}^X_1$ is therefore
\[	(\bar{\mu}^X_1)^T X >0	\hspace{0.2in}\text{or} \hspace{0.2in} u^T_1\Sigma_{YX}X>0\] 
Since $u^T_1\Sigma_{YX}X$ is a linear combination of $X$, it also has normal distribution 
\[	\mathcal{N} \left( 0, u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1\right)	\]
And $(Y,u^T_1\Sigma_{YX}X)$ also have multivariate normal distribution with mean $\mathbf{0}$ 
and covariance matrix
\[
\begin{bmatrix}
\Sigma_{YY} & \Sigma_{YX}\Sigma_{XY}u_1  \\
u^T_1\Sigma_{YX}\Sigma_{XY} &  u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1
\end{bmatrix}
\]
The conditional distribution of $Y|u^T_1\Sigma_{YX}X$ is also multivariate normal with mean 
\[	
\mu_{Y|u^T_1\Sigma_{YX}X } = \Sigma_{YX}\Sigma_{XY}u_1 (u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1)^{-1}u^T_1\Sigma_{YX}X	
\]
The $Y$ center for $u^T_1\Sigma_{YX}X>0$ is
\begin{align}
\hat{\mu}^Y_1 &= E(Y|u^T_1\Sigma_{YX}X>0) \nonumber \\ \nonumber
 		      & =  \Sigma_{YX}\Sigma_{XY}u_1 (u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1)^{-1} E(u^T_1\Sigma_{YX}X \mid u^T_1\Sigma_{YX}X>0) \\ \nonumber
\end{align}
Note that $u^T_1\Sigma_{YX}X$ has normal distribution $\mathcal{N} \left( 0, u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1\right)$, so
\[
E(u^T_1\Sigma_{YX}X \mid u^T_1\Sigma_{YX}X>0) = \sqrt{2/\pi}\cdot\sqrt{u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1}
\]
Therefore, we have the $Y$ center for $u^T_1\Sigma_{YX}X>0$ be
\begin{align*}
\hat{\mu}^Y_1 &= \sqrt{2/\pi}\cdot\sqrt{u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1} \hspace{0.1in} \Sigma_{YX}\Sigma_{XY}u_1 (u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1)^{-1} \\
&=\frac{\sqrt{2/\pi}}{\sqrt{u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1}} \Sigma_{YX}\Sigma_{XY}u_1
\end{align*}

Recall that $\bar{\mu}^Y_1 =\sqrt{2 \lambda_1/\pi}u_1$, to judge if $CV(2) > CV(1)$, one only need to compare the distance between  $\hat{\mu}^Y_1$  and  $\bar{\mu}^Y_1$ with distance between  $\hat{\mu}^Y_1$ and grand mean $0$. By variance and bias decomposition of prediction MSE, when variance is the same, only bias influence the MSE.

After some linear algebra manipulation, we get
$||\hat{\mu}^Y_1 - \bar{\mu}^Y_1||^2 > ||\hat{\mu}^Y_1||^2$ or $CV(2) > CV(1)$ iff
\[
 \frac{\sqrt{\lambda_1}}{2} > \frac{u^T_1\Sigma_{YX}\Sigma_{XY}u_1}{\sqrt{u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1}} 
\]
\end{proof}

\subsection{Two clusters in two dimensions}

We will now analyze a simple two-cluster setting, and derive conditions for
Gabriel cross-validation to correctly prefer $k=2$ clusters to $k=1$.

\begin{proposition}
Suppose that $\{(X_i,Y_i)\}_{i=1}^{n+m}$ is data from a single fold of Gabriel
cross-validation, where each $(X,Y)$ pair in $\R^2$ is an independent draw
from an equiprobable mixture of two multivariate normal distributions with
identity covariance. Suppose that the first mixture component has mean $(\muX,
\muY)$, and the second has mean $(-\muX, -\muY)$, where $\muX > 0$ and
$\muY > 0$.  If $1 + 2\Phi(\mu^Y)+ \frac{2\varphi(\mu^Y)}{\mu^Y} < 4\Phi(\mu^X)$, then $\CV(2) < \CV(1)$ with
probability tending to one as $m$ and $n$ increase.
\end{proposition}

\begin{proof}
There are two clusters $G_1$ and $G_2$, where observations from $G_1$ are distributed as
\[	\mathcal{N}\left( \begin{pmatrix} 
    \mu^X \\
    \mu^Y \\
  \end{pmatrix}, \mathbf{I} \right)	\]
and observations from $G_2$ are distributed as
\[	\mathcal{N}\left( \begin{pmatrix} 
    -\mu^X \\
    -\mu^Y \\
  \end{pmatrix}, \mathbf{I} \right)	\]
where $\mu^X_1 > 0$ and $\mu^Y_1 > 0$. Let $G_i$ be the true cluster where observation $i$ is generated from, by assumption
\[	P(G_i=G_1) = P(G_i=G_2) = 1/2	\]
After applying $k$-means on $\{ Y_i\}_{i=1}^{n}$ with $k=2$, if $n$ is large enough, we have the estimated centroids $\bar{\mu}^Y_1$ and $\bar{\mu}^Y_2$ be close to $E(Y \mid Y>0)$ and $\E(Y \mid Y < 0)$,
with errors will be of size $\OhP(n^{-1/2})$. Here
\begin{align}
E(Y \mid Y>0) &= E(Y_1 \mid Y_1 > 0) \cdot P(Y_1>0) + E(Y_2 \mid Y_2 > 0) \cdot P(Y_2>0) \nonumber \\ 
   &= 2\varphi(\mu^Y)+2\mu^Y\Phi(\mu^Y)-\mu^Y  \\  \nonumber 
\end{align}
where $Y_1 \sim N(\mu^Y, 1)$ and $Y_2 \sim N(-\mu^Y, 1)$, and we used Lemma~\ref{lem:truncated-normal-moments}
(Appendix~\ref{app:technical-lemmas}). Similarly, we have
\begin{align}
E(Y \mid Y<0) &= E(Y_1 \mid Y_1 < 0) \cdot P(Y_1<0) + E(Y_2 \mid Y_2 < 0) \cdot P(Y_2<0) \nonumber \\ 
   &= -2\varphi(\mu^Y)-2\mu^Y\Phi(\mu^Y)+\mu^Y  \\  \nonumber 
\end{align}
 where $\varphi()$ and $\Phi()$ are the standard normal probability and cumulative distribution function respectively.

Same as in single cluster case, the classification rule learned from
$\{(X_i, \hGY_i)\}_{i=1}^{n}$ variables will be determined according to
whether $X > 0$, with the decision boundary be at $0 + \OhP(n^{-1/2})$.
By symmetry, the CV error for points from $G_1$ is same as the points from $G_2$. Because $P(G=G_1) = P(G=G_2) =1/2$, the CV error can be calculated solely from $G_2$, that is

\begin{align*}
  \CV(2)
  &=
    \frac{1}{m}
    \sum_{i=n+1}^{n+m}
      \| (Y_i - \bmuY_1) 1\{\hGX_i = 1\} \|^2
      +
      \| (Y_i - \bmuY_2) 1\{\hGX_i = 2\} \|^2, \hspace{0.2in}Y\sim N(-\mu^Y,1)
\\
  &=
    \E[(Y - a)^2 1\{ X > 0\}] + \E[(Y + a)^2 1\{X < 0\}]
      + \OhP(m^{-1/2}) + \OhP(n^{-1/2})\\
   \end{align*}
   With 
   \begin{align}
 \E[(Y - a)^2 1\{ X > 0\}] + \E[(Y + a)^2 1\{X < 0\}]  =& P(\hGX_i = 1)\cdot E[(Y - a)^2] + P(\hGX_i = 2)\cdot E[(Y + a)^2] \nonumber \\  \nonumber
 =& [1-\Phi(\mu^X)][var(Y)+(-\mu^Y - a)^2]+\\ \nonumber
  & \Phi(\mu^X)[var(Y)+(-\mu^Y + a)^2]\\  \nonumber
  =& [1-\Phi(\mu^X)][1+(\mu^Y + a)^2]+\Phi(\mu^X)[1+(\mu^Y -a )^2]   \\  \nonumber
  =& 1+(\mu^Y + a)^2 -4a\Phi(\mu^X)\mu^Y  \nonumber
\end{align}
where $a$ is given by $(1)$.

When $k=1$, the result is straight forward since the estimated centroid will approximately equal to
$0$, with error of size $\OhP(n^{-1/2})$. The cross-validation error
will be
\[
  \CV(1) = \frac{1}{m} \sum_{i=n+1}^{n+m} \| Y_i - \bar Y_n \|^2
         = 1 + (\mu^Y)^2 + \OhP(m^{-1/2}) + \OhP(n^{-1/2}).
\]
So if we have $1 + 2\Phi(\mu^Y)+ \frac{2\varphi(\mu^Y)}{\mu^Y} < 4\Phi(\mu^X)$, we have $\CV(2) < \CV(1)$
\end{proof}

We confirm this result with a simulation.  We perform $10$ replicates for each pair of $(\mu^X, \mu^Y)$, where both $\mu^X$ and $\mu^Y$ take value on grid of $[0,3]$ with step $0.1$.  In each
replicate, we generate $20000$ observations from two multivariate normal distributions with identity covariance, where one has mean $(\muX, \muY)$ and the other one has mean $(-\muX, -\muY)$.
We perform a single $2 \times 2$ fold of Gabriel cross-validation and report the
times (out of $10$ replicates) when $k=2$ is selected by the algorithm in stead of $k=1$. Figure~\ref{fig:overlap-color_plot} shows the frequency $k=2$ is selected by the algorithm for each pair of $(\mu^X, \mu^Y)$. The dark spot means high number (close to $10$) is selected by the algorithm, which means algorithm very likely will pick $k=2$ over $k=1$ for the corresponding $(\mu^X, \mu^Y)$. While light spot means algorithm prefer $k=1$ for the corresponding value of $(\mu^X, \mu^Y)$. We can see the simulation result perfectly align with the theoretical curve (the black line), which separates the $k=2$ zone from the $k=1$ zone. It demonstrates that the Gabriel cross-validation works exactly as it suppose to under such setting. The position of dark spots shows that when the two clusters are reasonably apart (not overlapping too heavily) in both dimensions, the Gabriel cross-validation is asymptotically consistent.

\begin{figure}[H]
\centering
\includegraphics[width=3in]{demo/overlap/color_plot.pdf}
\caption{Number of times $k=2$ is selected out of $10$ replicates for each pair of $(\muX, \muY)$. The heat map shows the frequency $k=2$ is selected by the algorithm, with light means low number (of $k=2$) is selected and dark color means high number is selected. The black line is the theoretical curve, above which the algorithm suppose to pick $k=2$ and below which algorithm select $k=1$.  }
\label{fig:overlap-color_plot}
\end{figure}

\section{Adjusting for correlation}
\label{sec:corr-correct}

When the correlation between dimensions are high, the proposed Gabriel CV method tends to overestimate the number $k$. A simply remedy for the high correlation is available if we assume common covariance structure among the $k$ clusters. 
The procedure consists of four steps as follows:
\begin{enumerate}
	\item Apply Gabriel CV method on the original data $\dataX$, get estimated number of cluster $\hat{k}$
	\item Estimate the pooled covariance matrix $\hat{\Sigma}$ from the $\hat{k}$ clusters.
	\item Let $\hat{\Sigma} = \Gamma\Lambda\Gamma'$ be the Eigendecomposition of $\hat{\Sigma}$, we rescale and rotate the original data $\dataX$ to get $ \widetilde{\dataX} = \dataX\Gamma\Lambda^{-1/2}Q$, where $Q$ is a random  orthonormal rotation matrix. 
	\item Apply Gabriel CV method again on the transformed data $ \widetilde{\dataX}$ to estimate $k$. 
\end{enumerate}
Such correlation-correction idea is based on the assumption that each observation $x_i$ in $\dataX$ can be expressed as
\[	x_i = \mu(G_i)+\varepsilon_i \]
where $\mu(G_i)$ is the true cluster center for $x_i$ and $\varepsilon_i$ is noise with covariance matrix $\Sigma$. To reduce correlation between dimensions, we would like to transform $\Sigma$ into a diagonal matrix. A simple way to achieve this goal is multiplying $\dataX$ by $\Sigma^{-1/2}$, so the transformed $x_i$

\[	\Sigma^{-1/2} x_i = \Sigma^{-1/2} \mu(G_i)+\Sigma^{-1/2} \varepsilon_i \]
has identity covariance matrix $\mathbf{I}$. This is exactly what step $3$ does. However, before we can transform the data, we need to estimate the covariance matrix $\Sigma$. Step $1$ and $2$ estimate the covariance matrix by first applying Gabriel CV method on $\dataX$ to get preliminary estimate $\hat{k}$ and then estimate $\Sigma$ through

\[	\hat{\Sigma} = \frac{1}{n-\hat{k}} \sum^n_{i=1} (x_i-\hat{\mu}_i)(x_i-\hat{\mu}_i)^T	\]
where $\hat{\mu}_i$ is the center of the cluster which contains $x_i$ when applying $k$-means on $\dataX$ with parameter $\hat{k}$.

In practice, $\hat{\Sigma}$ wouldn't be the efficient estimator of $\Sigma$ since step $1$ may yield $\hat{k} > k$. Nevertheless, above procedure works well with correlated noise as we shall see in the simulation section.

\section{Performance in simulations}
\label{sec:simulations}

In this section, simulation is performed to evaluate the performance of our
proposed methods in locating the ``correct'' number of clusters. We compare
with a basket of existing methods including Gap statistics
\citep{tibshirani2001estimating}, Gaussian mixture model-based clustering
\citep{fraley2002model}, CH-index \citep{calinski1974dendrite}, Hartigan
statistics \citep{hartigan1975clustering}, Jump method
\citep{sugar2003finding}, Prediction strength \citep{tibshirani2005cluster},
Bootstrap stability \citep{fang2012selection} in following simulation
settings. All methods are executed with their default parameter settings. 
We select $p=q$ and $5-$fold cross-validation in row
($m=\frac{1}{4}n$) as default parameter setting for our proposed Gabriel
method. Note that set $p=q$ corresponding to $2-$fold cross-validation in
column. Wold method (detailed in Appendix $A$) and correlation-corrected 
Gabriel method (section $4.3$) are also included for comparison.

Note that in all settings, cluster centers are randomly generated from 
multivariate normal distribution $\mathcal{N}\left(\mathbf{0},\varsigma \mathbf{I}\right)$. 
All clusters are well-separated, i.e. no overlapping. In fact, any simulated 
clusters with minimum distance less than one unit was discarded, so there is
clear definition of true number of clusters. The parameters $\varsigma$ is chosen 
such that about half of the random realization were discarded. The idea is borrowed from
\cite{tibshirani2001estimating}. The proportion of each method successfully picks the correct
$k$ out of $100$ simulation trials is reported, along with its corresponding confident interval
by Wilson's method \citep{wilson1927probable}.

\begin{enumerate}

  \item \textbf{Correlation between dimensions} -- Six clusters in $10$ dimensions. 
  	Each cluster has $100$ or $50$
    multivariate normal observations with common covariance matrix $\Sigma$ 
    which has compound symmetric structure with $1$ in diagonal and $\rho$ 
    off diagonal. $\rho$ takes value in $\{0,0.1,...,0.9\}$.

	\begin{figure}[H]
	\centering
	\includegraphics[width=5.5in, height=3.3in]{New simulation/demo/bench/setting1/Facet.pdf}
	\label{fig:setting1}
	\end{figure}

  \item \textbf{Noise dimensions} --- Three clusters in $6$ dimensions. Each cluster has $1000$ or $500$
    mean zero multivariate normal observations with identity covariance matrix. 
    We add $P$ dimensions of noise to the data, which is randomly generated from uniform$[0,1]$. The noise
    dimension $P$ takes values in $\{0,6,...,54\}$.

	\begin{figure}[H]
	\centering
	\includegraphics[width=5.5in, height=3.3in]{New simulation/demo/bench/setting2/Facet.pdf}
	\label{fig:setting2}
	\end{figure}

  \item \textbf{High dimension} --- Eight clusters in $p$ dimensions, $p$ takes values 
  	in $\{10,20,...,100\}$. Each cluster has $100$ or $50$ mean zero multivariate normal 
  	observations with identity covariance matrix.

	\begin{figure}[H]
	\centering
	\includegraphics[width=5.5in, height=3.3in]{New simulation/demo/bench/setting3/Facet.pdf}
	\label{fig:setting3}
	\end{figure}

     \item \textbf{Variance heterogeneity} --- Three clusters in $20$ dimensions, each has
     $60$ observations in it. Observations are generated from 
     $\mathcal{N}\left(\mathbf{0},\sigma_1^2\mathbf{I}\right)$,
      $\mathcal{N}\left(\mathbf{0},\sigma_2^2\mathbf{I}\right)$ and
       $\mathcal{N}\left(\mathbf{0},\sigma_3^2\mathbf{I}\right)$ where 
       $\sigma_1^2 : \sigma_2^2: \sigma_3^2 = 1:\frac{1+R}{2}:R$. The maximum ratio $R$
       takes values in $\{1,5,10,...,45\}$

	\begin{figure}[H]
	\centering
	\includegraphics[width=5.5in, height=3.3in]{New simulation/demo/bench/setting4/Facet.pdf}
	\label{fig:setting4}
	\end{figure}

	\item \textbf{Heavy tail data} --- Five clusters in $15$ dimensions, each has
     $80$ observations in it. Observations have independent $t$ distribution in each dimension  
     with degree of freedom $\nu$, which takes value in $\{11,10,...,2\}$

	\begin{figure}[H]
	\centering
	\includegraphics[width=5.5in, height=3.3in]{New simulation/demo/bench/setting5/Facet.pdf}
	\label{fig:setting4}
	\end{figure}
\end{enumerate}

From the graph of setting $1$, we can see high correlation between dimensions cause problem for most 
existing methods as well as the proposed Gabriel CV method and Wold method. Higher correlation clearly 
degenerate most algorithms' ability to find the correct $k$ as data no longer spherical in $p$-dimension.
And the recent proposed methods (Gabriel CV method, Gap statistic, Jump method) seem to be more sensitive to
high correlation than those old methods such as CH-index and Hartigan statistics. The only two methods 
work well in high correlation are the Gaussian model-based BIC method \citep{fraley2002model} and the 
correlation-corrected Gabriel method.

Plot of setting $3$ shows some methods are insensitive to higher dimension such as Jump method and Gap 
statistics, while others deteriorate quickly with increasing dimension, most notably the Gaussian 
model-based BIC method. Note here the data are still generated from mixture of Gaussian. In contrast, our 
proposed Gabriel CV method as well as its correlation-corrected version actually working better in higher 
dimension.

Result of setting $4$ indicates most previously methods are sensitive to variance heterogeneity among 
clusters, most notably the Gap statistics and model-based BIC method. The proposed Gabriel CV method
and its correlation-corrected version consistently perform well in estimating $k$ and quite insensitive
to variance heterogeneity. The Wold method also performs well in this setting.

The last non-Gaussian setting examines methods' performance in heavy-tail data. With degree of freedom 
decreases, the tail becomes more flat and the Gaussian assumption becomes more inappropriate. It explains
the sharp plunge of Gaussian model-based BIC method in the plot. For most methods, their performances are 
relatively stable until the tail gets very heavy. In case where degree of freedom is $2$, Gap statistics 
and Jump method dived considerably compare to the Gabriel method and Wold method.

In conclusion, the proposed Gabriel CV method compares well with current state-of-the-art methods and 
is very robust against variance heterogeneity, high dimension and heavy-tail data. One weakness of the 
Gabriel CV method is that it's sensitive to the high correlation among dimensions. In such case, its 
correlation-corrected version will perform well under the assumption of common covariance structure.

\section{Empirical validation}
\label{sec:empirical-validation}

We also applied our proposed method to three real world data sets, two were
obtained from the University of California Irvine machine learning repository.
The first and third data sets are selected because there are clear number of
clusters in those two data sets. The second data set is used as a benchmark
data set since it was widely used in literature.

The first one is congress voting data which consists of voting records of
$98$th United States Congress, $2$nd session \citep{schlimmer1987concept}.
This data set includes votes for each of the U.S. House of Representatives
Congressmen on the $16$ key votes identified by the \textit{CQA}
(Congressional Quarterly Almanac). For each vote, each Congressman either vote
positively ``yea" (voted for/paired for/announced for),  negatively ``nay''
(voted against/paired against/announced against) or position unknown ``?". We
took out those records contain ``?" before comparing each algorithm. It
results in $232$ remaining records, with $124$ democrat and $108$ republican.

The second data set is the well-known Wisconsin breast cancer data set
\citep{mangasarian1990pattern}. After excluding the records with missing data,
this data set consists records of $683$ patients, each with measurements of
nine attributes of their biopsy specimens. It is known that there exist at
least two groups of patients: $444$ patients with benign specimens and $239$
patients with malignant specimens.

The third data set is gene expression data of $5$ types of brain tumours from
\cite{de2008clustering}, which contains $42$ observations including $10$
medulloblastomas, $10$ malignant gliomas, $10$ atypical teratoid/rhabdoid
tumours (AT/RTs), $8$ primitive neuroectodermal tumours (PNETs) and $4$ normal
cerebella. It was originally used by \cite{pomeroy2002prediction} to show that
medulloblastomas are molecularly distinct from other types of brain tumours.

Data was collected with Affymetrix microarrays, which estimates the number of
RNA copies found in the cell sample (frozen specimens). Data was preprocessed
as following: as a common procedure, all genes with expression level below
$10$ are set to a minimum level threshold of $10$. The maximum threshold is
set at $16,000$. Values below or above these thresholds are often unreliable.
To remove uninformative genes, for each gene $j$ (attribute), the mean $m_j$
among the samples was calculated. The $10\%$ largest and smallest values are
discarded in order to remove extreme values. Based on the mean, following
transformation is applied on every value $x_{ij}$ of sample $i$ and gene j \[
y_{ij} = \log_2 (x_{ij}/m_j)	\]

Genes with expression levels differing by at least $l$-flod in at least $c$
samples from their mean expression level across the sample were selected. The
parameter $l$ and $c$ were chosen so as to produce a filtered data set with at
least $10\%$ of the original number of genes \citep{de2008clustering}. Such
preprocessing procedure results in $1379$ genes selected. After filtering
step, data is restored in the original scale (i.e. $x_{ij}$). So the data set
is $42 \times 1379$, which serves as a high dimensional example.

\begin{table}[H]
\begin{center}
\captionsetup{justification=centering}
\caption{\label{table2} Number of clusters selected by each algorithm}
\begin{tabular}{lccccc}
    \hline                  
 & Congress Voting && Breast Cancer && Brain Tumours \\ \hline                    
CH-index & $2$ && $2$  && $2$    \\
Hartigan & $3$ && $3$  &&  $4$   \\
Jump & $10$ && $9$ && $1$  \\   
Prediction strength & $2$ &&  $2$ && $1$    \\
Bootstrap stability & $2$ &&  $2$ &&  $7$   \\ 
Gap & $8$ && $10$ && $10$  \\   
Gaussian-BIC & $2$ && $5$ && $2$  \\   
Gabriel & $2$  && $3$  && $5$  \\    
Gabriel-corr-correct & $2$  && $2$  && $5$  \\  
Wold & $2$ && $3$ && $4$ \\  \hline

\end{tabular}
\end{center}
\hspace{0.5in} \footnotesize {All the algorithms executed with their default parameter settings with $k$ ranges from $1$ to $10$}
\end{table}

Since most congressmen vote based on their parties' interest, two parties
(Democratic and Republican) represent two clusters in this data set. So the
optimal number should be two. Close inspection shows $k$-means with $k=2$
separates the two parties very well with the lowest miss-classification error
$10.43\%$ ($k=3$ has $14.78\%$). Our proposed Gabriel CV method, along with
several other algorithms correctly pick $k=2$ for this data set. Note that
CH-index and Bootstrap stability also return $k=2$, but $2$ is the lower bound
those methods can select for $k$. So it's not clear they actually choose $k=2$
or they hit the lower bound (they would pick $k=1$ if allowed).

For the breast cancer data, it's known to have $2$ clusters based on whether
it's benign specimens or malignant specimens. Although the proposed Gabriel
method picks $k=3$, its correlation-corrected version correctly picks $k=2$.
It shows the necessity to correct for the high correlation among data
features. However, for this data $k=3$ is also an acceptable answer, as
\cite{fujita2014non} noticed that the malign group is quite heterogeneous and
can be further clustered into at least two subgroups.

The brain tumours gene expression data is a typical high-dimension data with
$p > n$. It is clear that the correct number of clusters should be $5$, the
number of brain tumour types. Indeed, $k$-means results with $k=5$ show that
the $4$ normal human cerebella are well separated from brain tumours as they
form a single cluster.  Medulloblastomas, malignant gliomas and AT/RT tumours
also form $3$ respective clusters, which means they can be separated from each
other. Such results are similar to those found in
\cite{pomeroy2002prediction}, whose analysis is based on the first $3$
principle components of the original data.    From table \ref{table2} we can
see only our proposed Gabriel method (as well as its correlation-corrected
version) correctly selects $k=5$, underline how difficult it was to pick the
right $k$ when dimension is high.

\section{Application on Yeast data set}
\label{sec:application-yeast}

The yeast Saccharomyces cerevisiae data set was collected by \cite{cho1998genome} to study the characterization of the cell cycle related processes of the budding yeast S. cerevisiae. To obtain data, cdc28 cells were synchronized by raising temperature to $37^\circ$C in late G1 phase, and then reinitiated by changing cells to cooler environment ($25^\circ$C). Data was collected at $17$ time points at every $10$ minutes interval covering almost $2$ complete cell cycles. Oligonucleotide microarrays was used to query $6,220$ gene expression profiles at these $17$ time points. Besides the original paper, several other studies have analyzed this data.

\cite{tavazoie1999systematic} applied $k$-means clustering on the preprocessed data with $k=30$ and discovered several periodic clusters. The preprocessing procedure includes choosing the most variable $2,945$ geness among the total $6,220$. The normalized dispersion across time points (s.d/mean) was used as the metric of variation for selection. Data at time points $90$ and $100$ minutes were removed because they were deemed unreliable. It results a $2,945$ by $15$ data matrix. The last step of the preprocessing procedure was to variance normalize each gene across the $15$ time points, which is done by subtracting the mean and then dividing by the standard deviation. The result is each gene has mean zero and standard deviation equal to one across the time points.

\cite{dortet2008model} performed model-based clustering on the preprocessed data of \cite{tavazoie1999systematic} and  found $27$ clusters. They classified the $27$ clusters into $3$ major types: cyclic gene clusters, stress gene clusters and the clusters neither cyclically nor sensitive to the induced stress.

We also applied our proposed Gabriel method on the preprocessed data of \cite{tavazoie1999systematic}. The algorithm found $5$ clusters. Note here we cluster the genes as observations into clusters and the $15$ time points are considered attributes of each gene. Since the dimensions (time points) are likely to be correlated, we also applied the correlation-corrected version of Gabriel method on the data which also result $5$ clusters. This numbers is much smaller compare to those of \cite{tavazoie1999systematic} and \cite{dortet2008model}. We will focus our analysis on the $5$ clusters found by the proposed method, whose result is summarized on the bottom panel of Figure \ref{fig:5_clusters}, where each cluster is presented by the mean expression level of all the genes within that cluster over $15$ time points. Similar as \cite{tavazoie1999systematic}, we found the members of each cluster to be significantly enriched for genes function in similar biological process. Each gene is mapped with the Gene Ontology (GO) Annotation in Saccharomyces Genome Database (SGD) database, which provides Gene Ontology (GO) annotation of the entire genome using GO slims. GO slims are subsets of GO terms that give broad overview of ontology content and allow grouping of genes into broad categories, such as biological process, molecular function or cellular component. We choose to group genes in terms of biological process, which is a recognized series of events or molecular functions. There are total $103$ categories for biological process.

To identify biologically meaningful gene clusters from spurious ones, \cite{tavazoie1999systematic} introduced a method, enrichment analysis, to assess the statistical significance of the over-represented gene class in a cluster using hypergeometric distribution. Specifically, the $p$-value of observing $k$ genes from a particular category group within a cluster of size $m$ is given by:
\[	P = 1-\sum^{k-1}_{i=1} \frac{{{2945-n}\choose{m-i}}{{n}\choose{i}} }{ {{2945} \choose{m}} }\] 
where $n$ is the total genes in that group. It can be seen as testing the null hypothesis that genes are randomly  distributed among each cluster. The categories with the most significant grouping of gene inside each of the $5$ clusters are shown in Table \ref{table:enrich}. Only $p$-values less than $4.8\times10^{-4}$ ($0.05/103$) are reported here.

To better compare our results with those of \cite{tavazoie1999systematic}, we also prepared confusion Table \ref{table:confusion} and Figure \ref{fig:5_clusters}. The number in cell $(i,j)$ of Table \ref{table:confusion} shows the number of observations that belong to both the $i$th cluster of \cite{tavazoie1999systematic}'s and the $j$th cluster of ours. The plots in Figure \ref{fig:5_clusters} have similar meaning -- the mini plot on first row and the $4$th column is the mean gene expression level of observations belong to both the first cluster of \cite{tavazoie1999systematic}'s and the $4$th cluster of ours. We only plot the mean expression level if the number of observations in that cell is greater than $20$. Although \cite{tavazoie1999systematic} found $30$ clusters, not all of them were discussed in \cite{tavazoie1999systematic}. We only compare our results with those clusters that were profiled in \cite{tavazoie1999systematic}. The complete confusion matrix is in Appendix \ref{app:confusion}. The original result of \cite{tavazoie1999systematic} is obtained from site \url{http://arep.med.harvard.edu/network_discovery/}.

\begin{table}
\begin{center}
\captionsetup{justification=centering}
\caption{\label{table:enrich} Enrichment of clusters for genes within biological process categories}
\scalebox{0.8}{%
\begin{tabular}{cclcl}
\hline
   Cluster & Cluster size($m$) & Biological process category(total genes) & genes within category($k$) &$p$-value \\ \hline
 $1$ & $550$  &  response to oxidative stress ($55$)  &    $24$  & $1.5\times10^{-5}$ \\
     &     &  response to chemical ($213$)    &    $64$     & $2.2\times10^{-5}$\\
  $2$& $590$  &  mitochondrion organization ($159$) & $79$ &  $1.1\times10^{-16}$  \\  
  	 &     &  mitochondrial translation ($51$) & $28$ 	&  $2.9\times10^{-8}$  \\  
  	 &     &  generation of precursor metabolites and energy ($80$) &  $37$ & $7.3\times10^{-8}$   \\  
 $3$ & $654$ &  transcription from RNA polymerase II promoter ($214$) & $75$ &  $5.5\times10^{-6}$    \\  
  	 &     & mRNA processing ($67$) & $30$ 	& $2.7\times10^{-5}$    \\  
  	 &     & mitotic cell cycle ($183$)  & $63$ 	& $6.2\times10^{-5}$  \\ 
  $4$& $634$  & cytoplasmic translation ($134$)  &  $105$ &  $3.3\times10^{-47}$   \\  
     &     &  ribosomal subunit biogenesis $(138)$ & $73$ 	&  $7.7\times10^{-17}$  \\  
  	 &     &  rRNA processing ($131$) &  $61$	&  $5.7\times10^{-11}$  \\  
  	 &     &  ribosome assembly ($36$) &  $21$	&  $1.5\times10^{-6}$  \\  
$5$  &   $517$   &  chromosome segregation($106$) &  $53$      &  $6.0\times10^{-15}$ \\
     &     &  cellular response to DNA damage stimulus ($172$)  &  $71$  & $3.6\times10^{-14}$     \\  
  	 &     &  DNA repair ($147$)  &  $64$  & $3.7\times10^{-14}$     \\                 
 	 &     &  DNA replication ($78$)  & $42$   &  $1.8\times10^{-13}$   \\  
  	 &     &  mitotic cell cycle ($183$)  &  $70$  &  $4.8\times10^{-12}$   \\

\hline
\end{tabular}
}
\end{center}
\hspace{0.5in} \footnotesize {}
\end{table}

It turns out that each cluster has unique trend of expression level and contains genes that tend to participate in the common processes. Cluster $1$ has decreasing expression level with time. From Table \ref{table:enrich}, we can see it concentrates genes that somatize the cell stress, such as oxidative heat-induce proteins. It mainly consists of observations that did not get profiled in \cite{tavazoie1999systematic}, as we can see from Table \ref{table:confusion} and Figure \ref{fig:5_clusters}.

The mean gene expression level in Cluster $2$ decreases at the beginning and then increase. Our analysis found genes govern mitochondrial translation and mitochondrion organization are heavily concentrated in cluster $2$. Genes in \cite{tavazoie1999systematic}'s cluster $4$ and $8$ are heavily concentrated in this cluster, which also explains the coincidence between the enriched categories in our cluster $2$ and the functional groups in cluster $4$ and $8$ of    \cite{tavazoie1999systematic}.

Our cluster $3$ is a periodic cluster where one can see two gene expression cycles from Figure \ref{fig:5_clusters}. Indeed, cell cycle genes that related to budding and cell polarity are commonly found in cluster $3$. In addition, one can see from Table \ref{table:enrich} that genes govern RNA processing and transcription are highly represented in this cluster. Comparing cluster $3$ to the clusters found in \cite{tavazoie1999systematic} from Table \ref{table:confusion} and Figure \ref{fig:5_clusters}, one can see that it mainly consists of genes from cluster $3$, $7$ and $14$ of \cite{tavazoie1999systematic}. Note that the cluster $7$ and $14$ are both periodic clusters highlighted in \cite{tavazoie1999systematic}.

Cluster $4$ has increasing expression level with time. From Table \ref{table:enrich}, we can see genes related to cytoplasmic translation are highlighted in this cluster. Genes encode ribosomes are also overly presented in cluster $4$. This cluster corresponding to cluster $1$, the most notable functional grouping cluster of \cite{tavazoie1999systematic}. Indeed, one can see from Table \ref{table:confusion} that almost entire cluster $1$ of  \cite{tavazoie1999systematic} is contained in our cluster $4$.

One can easily see from Figure \ref{fig:5_clusters} that our cluster $5$ is also a periodic cluster. Enrichment analysis shows that genes participated in the cell-cycle phase related processes such as DNA replication and DNA repair are highly concentrated in this cluster. Table \ref{table:confusion} shows that the entire cluster $2$ of \cite{tavazoie1999systematic} is contained in our cluster $5$, along with some genes from cluster $14$. Note that both cluster $2$ and $14$ are profiled periodic clusters in \cite{tavazoie1999systematic}. We can also see from Figure \ref{fig:5_clusters} that our cluster $5$ groups together genes with similar cycle phase in \cite{tavazoie1999systematic}'s cluster $2$ and $14$.

In conclusion, our clusters group together genes that participate in the common biological processes. Our results are broadly align with the results in \cite{tavazoie1999systematic}, which clustered the data with $k=30$. In particular, the periodic cell cycle related genes are concentrated in cluster $5$ and cluster $3$, which consistent with the findings in \cite{tavazoie1999systematic} (cluster $2$, $7$ and $14$). Indeed, more than $70\%$ genes which function in mitotic cell cycle process are concentrated in these two clusters. The number of clusters found by our method is much smaller than the number of clusters used in \cite{tavazoie1999systematic} and \cite{dortet2008model}. However, both studies found that most of their clusters have no characteristics, suggesting overestimate of the underlying diversity of biological expression classes in the data set \citep{tavazoie1999systematic}. From Figure \ref{fig:5_clusters}, one can see clusters in \cite{tavazoie1999systematic} with similar behavior are merged in our cluster, which leads to smaller number of clusters. Our result is more parsimonious, while still able to identify the functional grouping of genes in clusters highlighted in previous studies.

	\begin{figure}[H]
		\centering
	\includegraphics[width=5.7in, height=5.7in]{wei/split_plots.pdf}
	\caption{Yeast data set mean expression profiles. The $5$ clusters found by Gabriel method are on the bottom; Clusters profiled in \cite{tavazoie1999systematic} are on the right.}
	\label{fig:5_clusters}
	\end{figure}

\begin{table}[H]
\caption{\label{table:confusion} \small{Confusion matrix comparing $5$ clusters to clusters profiled in \cite{tavazoie1999systematic}}}
\centering
\begin{tabular}{lcccccr}
  \hline
 & Cluster 1 & Cluster 2 & Cluster 3 & Cluster 4 & Cluster 5 & Total \\ 
  \hline
Cluster 1 & 0 & 0 & 1 & 161 & 2 & 164 \\ 
  Cluster 2 & 1 & 0 & 0 & 0 & 185 & 186 \\ 
  Cluster 3 & 0 & 0 & 91 & 11 & 2 & 104 \\ 
  Cluster 4 & 0 & 102 & 2 & 66 & 0 & 170 \\ 
  Cluster 7 & 1 & 10 & 83 & 7 & 0 & 101 \\ 
  Cluster 8 & 3 & 145 & 0 & 0 & 0 & 148 \\ 
  Cluster 14 & 0 & 1 & 29 & 6 & 38 & 74 \\ 
  Other & 545 & 332 & 448 & 383 & 290 & 1998 \\ 
  Total & 550 & 590 & 654 & 634 & 517 & 2945 \\ 
   \hline
\end{tabular}
\end{table}

\section{Discussion}
\label{sec:discussion}

In this paper, we proposed a novel approach to estimate the number of
clusters. The intuition behind our proposed methods is to transfer the
unsupervised learning problem into supervised learning problem via novel form
of cross validation. Such approach is quite different from previous methods which 
utilize the within/between cluster dispersion or stability criterion for selecting the 
optimal $k$. Our method utilizes the connection between different dimensions (columns)
of data through the uniqueness of each cluster center. We proved the self-consistency 
for our proposed Gabriel CV method as well as its asymptotic property with Gaussian noise,
and showed the robustness of our method by simulation. Our method has very good performance
in our limited simulation settings and real data application, and clearly the superior
one when data has heterogeneous variance or is heavy-tailed.

Besides no strong modeling assumption is required, our proposed method is robust against data set 
with variance heterogeneity, unequal number of observations, non-Gaussian noise and high-dimension. 
Such robustness is important in practice because for any given data, it's hard to tell whether
its clusters have different number of observations, is the variance equal for each cluster,
or what underlying noise distribution it has. The weakness of our proposed method is 
that its theory assumes only week correlation between ``predictor'' columns and ``response'' 
columns. In practice, many data sets don't exhibit high correlations between columns, where
our proposed method can be safely applied. In case the high correlation does exist, 
the correlation-corrected version of Gabriel CV method in section $4.3$ can be used if we can 
assume common covariance structure. Other procedure such as leave out most columns for clustering 
may also be used to reduce its effect. However, the theory for such procedure has not
been fully developed and it could be the future research topic.

Another situation where our proposed method cannot be directly used is when the clusters 
are non-convex. In such situation, $k$-means itself doesn't work well, for example two 
concentric circles share the same cluster center \citep{hastie2009elements}. However, it's 
possible that our proposed Gabriel CV method can be used on the transformed data set. In the
concentric circles case where spectral clustering is appropriate, our proposed method can be applied 
on the eigenvector subspace of the graph Laplacian matrix inside the spectral clustering algorithm
to find the optimal $k$. This can also be the future research topic.

\appendix

\section{Clustering scree plot}

\label{sec:elbow-fail}

The top row of Figure~\ref{fig:elbow}
displays an example where the elbow in $W_k$ corresponds to the true number $k
= 4$ of mixture components in the data-generating mechanism. The elbow
approach is simple and often performs well, but it requires subjective
judgment as to where the elbow is located, and, as the bottom row of
Figure~\ref{fig:elbow} demonstrates, the approach can easily fail.

\begin{figure}
\centering
\begin{minipage}{\linewidth}
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\linewidth]{demo/elbow/correct-data.pdf}
  \end{minipage}
  \hspace{0.05in}
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\linewidth]{demo/elbow/correct-withinss.pdf}
  \end{minipage}
\end{minipage}
\begin{minipage}{\linewidth}
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\linewidth]{demo/elbow/incorrect-data.pdf}
  \end{minipage}
  \hspace{0.05in}
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\linewidth]{demo/elbow/incorrect-withinss.pdf}
  \end{minipage}
\end{minipage}
\caption{Left panels show the $(X,Y)$ data points; right panels
  show the corresponding values of the within-cluster sum of squares $W_k$
plotted against the number of clusters, $k$.}
\label{fig:elbow}
\end{figure}

\section{Review of consistency results for clustering}

Some proposed methods provide theoretical consistency result for choosing $k$.
Assume data is a mixture of $G$ $p$-dimension clusters with equal prior,
identically distributed with common covariance matrix and finite fourth
moments in each dimension, \citet{sugar2003finding} shows the Jump method will
pick the correct $k$ under the conditions that cluster centers are
sufficiently separated and an appropriate transformation is used.
Specifically, let $\Delta\sqrt{p}$ denotes the minimum Euclidean distance
between cluster centers, \citet{sugar2003finding} shows that the jump will be
maximized at $k=G$ given that $\Delta > 6$ and the existence of a positive $Y$
such that
\[
  \left(\frac{p\Delta^2W}{9G}\right)^{-Y}
    + \left(
        W \left[\frac{2^{2H^*(X)}}{K^2_{max}2\pi e}\right]
        - (\frac{\Delta}{6})^2(1-W)
      \right)^{-Y} < 2
\]
and $\left(\frac{p\Delta^2W}{9G}\right)^{-Y} <1/2$, where $H^*(X)$ is the
minimum entropy of the cluster membership over each dimension and $W =1-
\frac{6^4V_{\mathbf{X}}}{(\Delta^2-36)^2}$ with $V_{\mathbf{X}} = \var\left(
  1/p ||\mathbf{X}-\mu_j||^2_{\Gamma^{-1}} \mid \mathbf{X}
\hspace{0.05in}\text{in jth cluster}\right)$

\cite{tibshirani2005cluster} proved that the prediction strength method is
consistent under the setting that the observations in $p$-dimension are
uniformly distributed within one unit from their respective cluster centers,
with minimum distance between the $G$ cluster centers is four unit. Under such
well-separated setting, they show that 
\[ ps(G) = 1+ o_p(1), \hspace{0.5in} \underset{k \geq G+1}{\text{sup}} ps(k)\leq \frac{2}{3}+ o_p(1) \]
therefore $\hat{k}$ is consistent in estimating $G$.  When data does come from
mixture of Gaussian distribution, the model-based method of
\citet{fraley2002model} is consistent in estimating $k$ in low dimension.

Let $\Psi$ denotes any given base-clustering algorithm (e.g $k$-means), and
$\Psi(Z,k)$ the clustering obtained by applying $\Psi$ on data $Z$ with
parameter $k$. Also let $k_0$ be the optimal number of cluster and
$d\{\Psi(Z_1,k),\Psi(Z_2,k)\}$ be the measure of distance between clustering
$\Psi(Z_1,k)$ and $\Psi(Z_2,k)$ where $Z_1$ and $Z_2$ are two independent
samples from population. \citet{wang2010consistent} shows that their proposed
method will consistently select $k = k_0$, given such $k_0$ exists and
$d\{\Psi(Z_1,k),\Psi(Z_2,k)\}$ converges to zero at proper rate $r_k$.
However, $k_0$ is defined as the most stable number for $\Psi$ among all the
possible $k$, instead of the number of components in mixture model or
well-separated compact regions commonly seen in literature. Therefore, their
consistent result actually is convergent result, i.e. the selected $k$
converges to its limit $k_0$ given such limit exists.

\section{Wold CV estimation} \label{app:foobar}
\begin{itemize}
	\item For each $k = 1,2,...,k_{max}$
	\begin{enumerate}
		\item Randomly draw some entries in $\mathbf{X}$ missing, keep those hold-out values in vector $V_{true}$
		\item Impute the missing values with column mean or $0$, denote the imputed data as $\mathbf{X}_{new}$
		\item Apply the iterative procedure below until converge or stopping criteria reached
		\begin{itemize}
			\item Apply $K$-mean on data set $\mathbf{X}_{new}$ with parameter $k$
			\item Substitute each observation in $\mathbf{X}_{new}$ by its nearest center, get new data $\mathbf{X}^c_{new}$ ($\mathbf{X}_{new}$ keep the same)
			\item Replace (impute) those imputed values in $\mathbf{X}_{new}$ with the corresponding entries in $\mathbf{X}^c_{new}$
			\item Calculate the difference between the old and newly imputed values, check whether or not they coincide (converge) 
		\end{itemize}
		\item Obtain the last imputed entry values of converged $\mathbf{X}_{new}$, denote it by $V_{converge}$
		\item Calculate the prediction error  $Error_k = ||V_{true} - V_{converge}||^2$ 
	\end{enumerate}
	\item For each CV folder, repeat above procedure and obtain the $Error_k$ for each $k$
	\item Average $Error_k$ across all folders for each $k$, and then select the $k$ corresponding to the minimum average $Error_k$ 
\end{itemize}

\section{Confusion matrix} \label{app:confusion}

\begin{table}
%\captionsetup{justification=centering}
\caption{\label{table:confusion-complete} Complete confusion matrix comparing $5$ clusters to $30$ clusters }
\begin{tabular}{lccccc|c}
  \hline
 & Cluster 1 & Cluster 2 & Cluster 3 & Cluster 4 & Cluster 5 & Total \\ 
  \hline
  Cluster 1 & 0 & 0 & 1 & 161 & 2 & 164 \\ 
  Cluster 2 & 1 & 0 & 0 & 0 & 185 & 186 \\ 
  Cluster 3 & 0 & 0 & 91 & 11 & 2 & 104 \\ 
  Cluster 4 & 0 & 102 & 2 & 66 & 0 & 170 \\ 
  Cluster 5 & 152 & 0 & 0 & 0 & 0 & 152 \\ 
  Cluster 6 & 1 & 2 & 12 & 85 & 4 & 104 \\ 
  Cluster 7 & 1 & 10 & 83 & 7 & 0 & 101 \\ 
  Cluster 8 & 3 & 145 & 0 & 0 & 0 & 148 \\ 
  Cluster 9 & 142 & 0 & 0 & 0 & 5 & 147 \\ 
  Cluster 10 & 12 & 27 & 24 & 20 & 6 & 89 \\ 
  Cluster 11 & 0 & 5 & 81 & 8 & 0 & 94 \\ 
  Cluster 12 & 3 & 68 & 5 & 3 & 1 & 80 \\ 
  Cluster 13 & 0 & 1 & 12 & 81 & 5 & 99 \\ 
  Cluster 14 & 0 & 1 & 29 & 6 & 38 & 74 \\ 
  Cluster 15 & 60 & 53 & 2 & 0 & 0 & 115 \\ 
  Cluster 16 & 7 & 0 & 0 & 9 & 83 & 99 \\ 
  Cluster 17 & 9 & 0 & 72 & 1 & 1 & 83 \\ 
  Cluster 18 & 3 & 16 & 10 & 67 & 5 & 101 \\ 
  Cluster 19 & 62 & 2 & 0 & 1 & 8 & 73 \\ 
  Cluster 20 & 11 & 0 & 6 & 2 & 65 & 84 \\ 
  Cluster 21 & 1 & 13 & 6 & 44 & 6 & 70 \\ 
  Cluster 22 & 33 & 15 & 28 & 9 & 0 & 85 \\ 
  Cluster 23 & 2 & 18 & 2 & 25 & 22 & 69 \\ 
  Cluster 24 & 1 & 38 & 45 & 1 & 0 & 85 \\ 
  Cluster 25 & 2 & 57 & 16 & 1 & 0 & 76 \\ 
  Cluster 26 & 5 & 3 & 23 & 19 & 0 & 50 \\ 
  Cluster 27 & 9 & 0 & 30 & 5 & 20 & 64 \\ 
  Cluster 28 & 0 & 7 & 23 & 0 & 38 & 68 \\ 
  Cluster 29 & 27 & 2 & 10 & 1 & 11 & 51 \\ 
  Cluster 30 & 3 & 5 & 41 & 1 & 10 & 60 \\ \hline
  Total & 550 & 590 & 654 & 634 & 517 & 2945 \\ \hline
\end{tabular}
\end{table}

\bibliographystyle{apalike}

\bibliography{references}
\end{document}