### 2019-01-10 JCGS-R2/response/response.tex
start response

\documentclass[11pt]{article}

\usepackage{amsmath}
\usepackage{framed}
\usepackage{fullpage}
\usepackage{hyperref} % must be last usepackage
\usepackage{color}
\usepackage[authoryear]{natbib}

\newcommand{\hilight}[1]{\colorbox{yellow}{#1}}
\newcommand{\todo}{\hilight{TODO}}

\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\let\PBS=\PreserveBackslash

\title{``Estimating the number of clusters using cross-validation''
Summary of changes to the revised manuscript}
\author{Wei Fu \and Patrick O. Perry}

\begin{document}

\maketitle
\setlength{\parindent}{0ex}
\setlength{\parskip}{1ex}
\vspace{2ex}

We thank the editors and reviewers for their time reading our submission, and
for providing many helpful suggestions for improvement.  We also thank you for
the opportunity to revise and resubmit our manuscript.  We have revised our
manuscript carefully taking into account your comments.

Detailed responses to reviewer comments follow below.

\section*{Response to Editor}

\begin{itemize}

\item \textit{Fix the figure captions.}

\textbf{Response:} Per your instruction, all figure captions now include
(1) what is the plot about, (2) specific details of plot, like what type of
display and how variables are mapped, (3) the most important thing that the
reader should learn.

\todo

\item \textit{Cite R packages.}

\textbf{Response:} We added citations for all packages used in our work: \todo

% You need to cite R and any packages used in your work, like ggplot2.

\item \textit{Reprodicibility.}

\textbf{Response:} \todo

% Your code is quite a mess. It is not clear to me how to reproduce the results
% you have reported. There is code in two different tar.gz files. Which is the
% code to use? In each there are several folders. This all needs to be
% organised so that it is easy for the reader to produce the results that you
% have reported in the paper.

\end{itemize}

\section*{Response to Associate Editor}

\begin{itemize}

\item \textit{Compare to the slope method}

\todo

\end{itemize}

\section*{Response to Reviewer 1}

(No further comments.)

\section*{Response to Reviewer 2}

\begin{itemize}

\item \textit{Bottom of page 4. The authors need to clarify the notation}

\textbf{Response:} \todo

%The fold numbers r and r_i are not well-defined. Please maintain the
%notational consistency. You defined only "r_i". What do you mean by: 'For each
%"fold" r'?

\item \textit{Page 6. Clarify that $n \gg K$, $m \gg L$.}

\textbf{Response:} \todo

%By my understanding, the authors must assume further that
%
%n  >> K
%m >> K	
%
%That is, they need to ensure that they are choosing n and m larger that K.
%This limitation should be clarified to the reader.

\item \textit{Section 2.2. Clarify.}

\textbf{Response:} \todo

%The authors need to clarify if and how the methods are applied
%to a non-regression setting.

\item \textit{Section 3. Note that slope is self consistent.}

\textbf{Response:} \todo

%Section 3. I am very confident that the slope statistics (Fujita et al) is
%also self-consistent under the absence of noise. This follows basically from
%the properties of finite differences.  On the other hand, the original
%Silhouette method is not self-consistent. The authors could explore this
%easily in this section in one paragraph.

\item \textit{Proposition 1 improvements.}

\textbf{Response:} \todo

We chose to keep the proof in the main text, but if there are space
constraints we will take your suggestion and move the proof to an appendix.

% 5. Proposition 1. 
% 
% 5.1 In order to check condition (iv), it seems the user must know something about K.
% 
% 5.2 Should it be CV(k)>CV(K)?
% 
% 5.3 My feeling is that once the reader fully understand the ingredients of the
% procedure, Proposition one renders trivial.  The authors could provide direct
% arguments for that. This seems to follow straightforward from the properties
% of estimation in a regression setting.

\item {\textit Proposition 2 improvements.}

\textbf{Response:} \todo

% Please, define the indicator function.

\item \textit{Page 8, line 14. Generalization.}

\textbf{Response:} \todo

We clarified what we meant by saying the proof ``generalizes to non-normal
distributions easily''. You rightly point out that it will not generalize to
all possible situations, but only to distributions with finite moments and
cases where both clusters have the same distribution.

\item \textit{Figure 2. Please define correctly the x,y axis}

\textbf{Response:} \todo

\item \textit{Page 20, first paragraph.}

\textbf{Response:} Your comment was that ``Proposition 2 only shows that if
$|\rho|<0.5$ the method works correctly.''Did you mean page 14 instead?  We
think you may have meant that our statement in the first paragraph of  page
14, ``Proposition~2 tells us that Gabriel cross-validation fails when there is
strong correlation between the variables'' is too strong.

\item \textit{Compare to standardized Euclidean norm $k$-means.}

\textbf{Response:} \todo

%10. Page 22, line 3. Maybe the standardized Euclidean norm in the k-means
%algorithm could produce good results.

\item \textit{Compare to slope}

\textbf{Response:} \todo

% 11. Page 23, line 35. Under normal distributions, the slope statistics
% proposed by Fujita et al. also works very good if the clusters are well
% separated. There are two R packages for computing the slope statistics: 
% 
% https://www.rdocumentation.org/packages/anocva/versions/0.1.1/topics/optimalSlope
% 
% and:
% 
% https://cran.r-project.org/web/packages/cstab/cstab.pdf
% 
% 
% 12. Page 25, line 31. The Slope statistic also works well under correlation.
% See the papers simulations (section 5).  Moreover, in Fujita et al., it is
% showed that some of the methods used here have bad performances under some
% scenarios which slope statistic was very good (high dimensional data, t and
% exponential distributions for each cluster, correlated items, Gaussian and
% uniform, heterogeneity).

\end{itemize}

\end{document}