### 2019-06-29 cover/cover.tex
update with local files

\documentclass[12pt,letterpaper]{article}

\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{slantsc}

\def\name{Patrick O. Perry}

\hypersetup{
  colorlinks = true,
  urlcolor = black,
  pdfauthor = {\name},
  pdfkeywords = {},
  pdftitle = {},
  pdfsubject = {},
  pdfpagemode = UseNone
}

\geometry{
  body={6.25in, 9.0in},
  left=1.25in,
  top=1.0in
}

\begin{document}
\pagenumbering{gobble}

\noindent
\begin{minipage}[t]{2in}
  \vspace{0pt}
  \includegraphics[width=1.42in]{logo.png}
\end{minipage}
\hfill
\begin{minipage}[t]{3in}
  \vspace{0pt}\raggedright
  {\footnotesize%
  {\scshape%
    Department~of~Information,~Operations,~\& \\
    Management Sciences \\
  }
  \vspace{\baselineskip}
  {\scshape%
    Henry Kaufman Management Center \\
    44 West 4\textsuperscript{th} Street, Suite 8-63 \\
    New York, NY 10012-1125 \\
    (212) 998-0194 \hspace{2ex} Fax: (212) 995-4227 \\
    Email:} pperry@stern.nyu.edu \\
  \vspace{\baselineskip}
  {\scshape {\bfseries Patrick O.\ Perry}} \\
  {\slshape%
    Assistant~Professor~of~Information,~Operations,~\& \\
    Management Sciences
  }
  }
\end{minipage}

\vspace{1\baselineskip}

\vspace{\stretch{1}}

\begin{flushright}
\today
\end{flushright}

\noindent
Dear Prof. Dianne Cook and JCGS Associate Editors,

\vspace{1\baselineskip}

Wei Fu and I are submitting revision of Manuscript ID JCGS-17-028. We have
corrected the mistakes and added discussion of the points raised by the AE and
the reviewers. The Associate Editor and Reviewer 1 mostly requested more
discussion of some of the difficult issues in the paper; we have added these
as suggested. Reviewer 2 requested that we generalize our Propositions by
removing normality and dimensionality assumptions; we have done this, either
with a new proof in the appendix or with discussion of the assumption after
the relevant proof. (We kept the assumptions in the main text as-is for the
sake of clarity of presentation.) Reviewer~2 also asked for the results in
table form rather than as figures. We have added these tables in a
supplementary appendix.

Not counting the supplementary appendix for the tables, the revision added
8 pages of material to the manuscript.

\vspace{1\baselineskip}

\noindent
Regards,

\vspace{4\baselineskip}
\noindent
Patrick Perry

\vspace{\stretch{1}}

\end{document}