### 2019-06-29 JCGS-R3/appendix.tex
fix margin, split out appendix

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{float}
%\usepackage{footnote}
%\usepackage[bottom]{footmisc}
\usepackage[singlelinecheck=false]{caption}

\usepackage{graphicx,psfrag,epsf}
\usepackage[space]{grffile}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{1}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{claim}{Claim}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}

% operators
\DeclareMathOperator*{\argmin}{arg\,min}
\newcommand{\CV}{\operatorname{CV}}
\newcommand{\PE}{\operatorname{PE}}
\newcommand{\E}{\operatorname{E}}
\DeclareMathOperator*{\cov}{cov}
\DeclareMathOperator*{\cor}{cor}
\DeclareMathOperator*{\tr}{tr}
\DeclareMathOperator*{\var}{var}
\newcommand{\T}{T}

% convergence
\newcommand{\toas}{\overset{\mathit{a.s.}}{\to}}

\newcommand{\Oh}{O}
\newcommand{\OhP}{O_p}

% sets
\newcommand{\R}{\mathbb{R}}
\newcommand{\sA}{\mathcal{A}}
\newcommand{\sbA}{\mathcal{\bar A}}
\newcommand{\sC}{\mathcal{C}}
\newcommand{\sE}{\mathcal{E}}
\newcommand{\sN}{\mathcal{N}}
\newcommand{\sX}{\mathcal{X}}

% scalars
\newcommand{\bpi}{\bar \pi}

% vectors
\newcommand{\muX}{\mu^{X}}
\newcommand{\muY}{\mu^{Y}}
\newcommand{\bmuX}{\bar \mu^{X}}
\newcommand{\bmuY}{\bar \mu^{Y}}
\newcommand{\hmuY}{\hat \mu^{Y}}

% matrices
\newcommand{\dataX}{\mathfrak{X}}
\newcommand{\SigmaY}{\Sigma^Y}
\newcommand{\Xtrain}{X_{\text{train}}}
\newcommand{\Ytrain}{Y_{\text{train}}}
\newcommand{\Xtest}{X_{\text{test}}}
\newcommand{\Ytest}{Y_{\text{test}}}

% class labels
\newcommand{\hGX}{\hat G^{X}}
\newcommand{\hGY}{\hat G^{Y}}

\begin{document}

%\bibliographystyle{natbib}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if1\blind
{
  \title{\bf Supplement to ``Estimating the number of clusters using
cross-validation''}
  \author{Wei Fu and Patrick O. Perry \\
  Stern School of Business, New York University}
  \maketitle
} \fi

\if0\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Estimating the number of clusters using cross-validation}
\end{center}
  \medskip
} \fi

\bigskip

\spacingset{1.45} % DON'T change the spacing!

\section{Clustering scree plot examples}

\label{sec:elbow-fail}

The top row of Figure~\ref{fig:elbow}
displays an example where the elbow in $W_k$ corresponds to the true number $k
= 4$ of mixture components in the data-generating mechanism. The elbow
approach is simple and often performs well, but it requires subjective
judgment as to where the elbow is located, and, as the bottom row of
Figure~\ref{fig:elbow} demonstrates, the approach can easily fail.

\begin{figure}
\centering
\begin{minipage}{\linewidth}
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\linewidth]{main_code/demo/elbow/correct-data.pdf}
  \end{minipage}
  \hspace{0.05in}
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\linewidth]{main_code/demo/elbow/correct-withinss.pdf}
  \end{minipage}
\end{minipage}
\begin{minipage}{\linewidth}
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\linewidth]{main_code/demo/elbow/incorrect-data.pdf}
  \end{minipage}
  \hspace{0.05in}
  \begin{minipage}{0.45\linewidth}
    \includegraphics[width=\linewidth]{main_code/demo/elbow/incorrect-withinss.pdf}
  \end{minipage}
\end{minipage}
\caption{Scree plots for two data sets. Left panels show the sets of
    two-dimensional data points, generated from four clusters, with plotting
    symbol indicating the generated cluster. Right panels show the
    corresponding values of the within-cluster sum of squares $W_k$
    plotted against the number of clusters, $k$. The scree plot identifies
    the correct number of clusters in the top row, but fails in the bottom
    row.}
\label{fig:elbow}
\end{figure}

%% \section{Review of consistency results for clustering}
%% 
%% Some proposed methods provide theoretical consistency result for choosing $k$.
%% Assume data is a mixture of $G$ $p$-dimension clusters with equal prior,
%% identically distributed with common covariance matrix and finite fourth
%% moments in each dimension, \citet{sugar2003finding} shows the Jump method will
%% pick the correct $k$ under the conditions that cluster centers are
%% sufficiently separated and an appropriate transformation is used.
%% Specifically, let $\Delta\sqrt{p}$ denotes the minimum Euclidean distance
%% between cluster centers, \citet{sugar2003finding} shows that the jump will be
%% maximized at $k=G$ given that $\Delta > 6$ and the existence of a positive $Y$
%% such that
%% \[
%%   \left(\frac{p\Delta^2W}{9G}\right)^{-Y}
%%     + \left(
%%         W \left[\frac{2^{2H^*(X)}}{K^2_{max}2\pi e}\right]
%%         - (\frac{\Delta}{6})^2(1-W)
%%       \right)^{-Y} < 2
%% \]
%% and $\left(\frac{p\Delta^2W}{9G}\right)^{-Y} <1/2$, where $H^*(X)$ is the
%% minimum entropy of the cluster membership over each dimension and $W =1-
%% \frac{6^4V_{\mathbf{X}}}{(\Delta^2-36)^2}$ with $V_{\mathbf{X}} = \var\left(
%%   1/p ||\mathbf{X}-\mu_j||^2_{\Gamma^{-1}} \mid \mathbf{X}
%% \hspace{0.05in}\text{in jth cluster}\right)$
%% 
%% \cite{tibshirani2005cluster} proved that the prediction strength method is
%% consistent under the setting that the observations in $p$-dimension are
%% uniformly distributed within one unit from their respective cluster centers,
%% with minimum distance between the $G$ cluster centers is four unit. Under such
%% well-separated setting, they show that 
%% \[ ps(G) = 1+ o_p(1), \hspace{0.5in} \underset{k \geq G+1}{\text{sup}} ps(k)\leq \frac{2}{3}+ o_p(1) \]
%% therefore $\hat{k}$ is consistent in estimating $G$.  When data does come from
%% mixture of Gaussian distribution, the model-based method of
%% \citet{fraley2002model} is consistent in estimating $k$ in low dimension.
%% 
%% Let $\Psi$ denotes any given base-clustering algorithm (e.g $k$-means), and
%% $\Psi(Z,k)$ the clustering obtained by applying $\Psi$ on data $Z$ with
%% parameter $k$. Also let $k_0$ be the optimal number of cluster and
%% $d\{\Psi(Z_1,k),\Psi(Z_2,k)\}$ be the measure of distance between clustering
%% $\Psi(Z_1,k)$ and $\Psi(Z_2,k)$ where $Z_1$ and $Z_2$ are two independent
%% samples from population. \citet{wang2010consistent} shows that their proposed
%% method will consistently select $k = k_0$, given such $k_0$ exists and
%% $d\{\Psi(Z_1,k),\Psi(Z_2,k)\}$ converges to zero at proper rate $r_k$.
%% However, $k_0$ is defined as the most stable number for $\Psi$ among all the
%% possible $k$, instead of the number of components in mixture model or
%% well-separated compact regions commonly seen in literature. Therefore, their
%% consistent result actually is convergent result, i.e. the selected $k$
%% converges to its limit $k_0$ given such limit exists.

\section{Analysis of single cluster in more than two dimensions}
\label{sec:single-general-dim}

\begin{proposition}\label{prop:single-general-dim}
Suppose that $\{ (X_i, Y_i) \}_{i=1}^{n + m}$ is data from a single fold
of Gabriel cross-validation, where each $(X,Y)$ pair in $\R^{p+q}$ is an
independent draw from a mean-zero multivariate normal distribution with 
covariance matrix $\Sigma = \left( \begin{smallmatrix} \Sigma_{XX} & \Sigma_{XY} \\ 
 \Sigma_{YX} & \Sigma_{YY} \end{smallmatrix}\right)$, with $\Sigma_{YY}$ has leading 
eigenvalue $\lambda_1$ and corresponding eigenvector $u_1$. In this case, the data are drawn
from a single cluster; the true number of clusters is~$1$.  If  $\frac{\sqrt{\lambda_1}}{2}
 > \frac{u^T_1\Sigma_{YX}\Sigma_{XY}u_1}{\sqrt{u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1}}$,
then $\CV(1) < \CV(2)$ with probability tending to one as $m$ and $n$ increase.
\end{proposition}

\begin{proof}
Let $X$ and $Y$ be jointly multivariate normal distributed with mean $\mathbf{0}$ and covariance matrix $\Sigma$, i.e.
\[	(X,Y) \sim \mathcal{N} \left( \mathbf{0}, \Sigma\right)	\]
where $\Sigma=\begin{bmatrix} \Sigma_{XX} & \Sigma_{XY} \\  \Sigma_{YX} & \Sigma_{YY} \end{bmatrix}$.

Let $\Sigma_{YY} = U \Lambda U^T$  be the eigendecomposition of $\Sigma_{YY}$,
with leading eigenvalue $\lambda_1$ and corresponding eigenvector $u_1$. Then
the centroid of $k$-means applying on $(y_1,..,y_n)$ is on the first principal
component
of $Y$,\[	E(u^T_1 Y|u^T_1 Y>0) = \bar{\mu}^Y_1 =\sqrt{2 \lambda_1/\pi}u_1\] and 
\[	E(u^T_1 Y|u^T_1 Y<0) = \bar{\mu}^Y_2 =-\sqrt{2 \lambda_1/\pi}u_1\]
 where $u^T_1 Y \sim \mathcal{N}(0,\lambda_1)$.

To compute $\bar{\mu}^X_1 = E(X|u^T_1 Y>0)$, we need to know the conditional distribution $X|u^T_1 Y$. Since $(X,Y)$ has multivariate normal distribution, $(X,u^T_1 Y)$ also has a multivariate normal distribution with mean $\mathbf{0}$ and covariance matrix
$$\Sigma_{X,u^T_1 Y}=\begin{bmatrix} \Sigma_{XX} & \Sigma_{XY} u_1 \\  u^T_1 \Sigma_{YX} & \lambda_1 \end{bmatrix}$$
The conditional distribution $X|u^T_1 Y$ is hence normal with mean
 $$\mu_{X|u^T_1 Y} = \Sigma_{XY} u_1 \lambda^{-1}_1 u^T_1 Y $$
Therefore, 
\begin{align}
\bar{\mu}^X_1 &= E(X \mid u^T_1 Y>0) \nonumber \\ \nonumber
 			  &= E\left(E[X \mid u^T_1 Y] \mid u^T_1 Y>0\right) \\ \nonumber
 			  &=  E\left(\Sigma_{XY} u_1 \lambda^{-1}_1 u^T_1 Y \mid u^T_1 Y>0\right)\\ \nonumber
 			  &= \lambda^{-1}_1 \Sigma_{XY}u_1 E(u^T_1 Y \mid u^T_1 Y>0) \\ \nonumber
 			  &= \lambda^{-1}_1 \Sigma_{XY}u_1 \sqrt{2 \lambda_1/\pi} \\ \nonumber
      &= \sqrt{2 / (\lambda_1 \pi)} \Sigma_{XY}u_1
\end{align}
Similar calculation yields $\bar{\mu}^X_2 = -\sqrt{2 / (\lambda_1 \pi)} \Sigma_{XY}u_1$.
The decision rule to classify any observed value of $X$ to $\bar{\mu}^X_1$ is therefore
\[	(\bar{\mu}^X_1)^T X >0	\hspace{0.2in}\text{or} \hspace{0.2in} u^T_1\Sigma_{YX}X>0\] 
Since $u^T_1\Sigma_{YX}X$ is a linear combination of $X$, it also has normal distribution 
\[	\mathcal{N} \left( 0, u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1\right)	\]
And $(Y,u^T_1\Sigma_{YX}X)$ also have multivariate normal distribution with mean $\mathbf{0}$ 
and covariance matrix
\[
\begin{bmatrix}
\Sigma_{YY} & \Sigma_{YX}\Sigma_{XY}u_1  \\
u^T_1\Sigma_{YX}\Sigma_{XY} &  u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1
\end{bmatrix}
\]
The conditional distribution of $Y|u^T_1\Sigma_{YX}X$ is also multivariate normal with mean 
\[	
\mu_{Y|u^T_1\Sigma_{YX}X } = \Sigma_{YX}\Sigma_{XY}u_1 (u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1)^{-1}u^T_1\Sigma_{YX}X	
\]
The $Y$ center for $u^T_1\Sigma_{YX}X>0$ is
\begin{align}
\hat{\mu}^Y_1 &= E(Y|u^T_1\Sigma_{YX}X>0) \nonumber \\ \nonumber
 		      & =  \Sigma_{YX}\Sigma_{XY}u_1 (u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1)^{-1} E(u^T_1\Sigma_{YX}X \mid u^T_1\Sigma_{YX}X>0) \\ \nonumber
\end{align}
Note that $u^T_1\Sigma_{YX}X$ has normal distribution $\mathcal{N} \left( 0, u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1\right)$, so
\[
E(u^T_1\Sigma_{YX}X \mid u^T_1\Sigma_{YX}X>0) = \sqrt{2/\pi}\cdot\sqrt{u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1}
\]
Therefore, we have the $Y$ center for $u^T_1\Sigma_{YX}X>0$ be
\begin{align*}
\hat{\mu}^Y_1 &= \sqrt{2/\pi}\cdot\sqrt{u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1} \hspace{0.1in} \Sigma_{YX}\Sigma_{XY}u_1 (u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1)^{-1} \\
&=\frac{\sqrt{2/\pi}}{\sqrt{u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1}} \Sigma_{YX}\Sigma_{XY}u_1
\end{align*}

Recall that $\bar{\mu}^Y_1 =\sqrt{2 \lambda_1/\pi}u_1$, to judge if $\CV(2) >
\CV(1)$, one only need to compare the distance between  $\hat{\mu}^Y_1$  and
$\bar{\mu}^Y_1$ with distance between  $\hat{\mu}^Y_1$ and grand mean $0$. By
the variance and bias decomposition of prediction MSE, when variance is the
same, only bias influences the MSE.

After some linear algebra manipulation, we get
$\|\hat{\mu}^Y_1 - \bar{\mu}^Y_1\|^2 > \|\hat{\mu}^Y_1\|^2$ or $\CV(2) >
\CV(1)$ if and only if
\[
 \frac{\sqrt{\lambda_1}}{2} > \frac{u^T_1\Sigma_{YX}\Sigma_{XY}u_1}{\sqrt{u^T_1\Sigma_{YX} \Sigma_{XX} \Sigma_{XY} u_1}} 
\]
\end{proof}

The condition that ensures $\CV(1) < \CV(2)$ is tight but difficult to
understand intuitively. We make understand this condition better by
noting that it is equivalent to a condition on the correlation between the
two linear combinations of $X$ and $Y$ variables:
\begin{equation}\label{eqn:clust-anyd}
    \cor(u_1^\T \Sigma_{YX} X, u_1^\T Y) < 1/2.
\end{equation}
A stronger condition that is easier to understand, then, is that the canonical
correlation between $X$ and $Y$ is below $1/2$; when the latter is true the
former follows and $\CV(1) < \CV(2)$. A necessary but not sufficient
guarantee of the stronger condition is that all correlations between variables
are below $1/2$.

One might object to the condition~\eqref{eqn:clust-anyd} being stated in terms
of a particular $(X, Y)$ split of the data.  In practice, one might want the
condition to hold for all possible splits of the data. A sufficient guarantee
that~\eqref{eqn:clust-anyd} holds for any $(X, Y)$ split of the data is that
the data covariance matrix $\Sigma$ is diagonal. The ad-hoc adjustment we
propose in Sec.~\ref{sec:corr-correct} attempts to transform the data to this
form.

\section{Analysis of two clusters in more than two dimensions}
\label{app:twoclust-anyd}

\begin{proposition}\label{prop:twoclust-anyd}
Suppose that $\{(X_i,Y_i)\}_{i=1}^{n+m}$ is data from a single fold of Gabriel
    cross-validation, where each $(X,Y)$ pair in with $X \in \mathbb{R}^P$ and $Y \in
    \mathbb{R}^Q$ is an independent draw
from an equiprobable mixture of two multivariate normal distributions with
identity covariance. Suppose that the first mixture component has mean
$\mu = (\muX, \muY)$ and the second has mean $-\mu = (-\muX, -\muY)$,
where $\muX \geq 0$ and $\muY \geq 0$. 
If the cluster centers are well separated, specifically such that
\(
  2 \varphi(\muY) + \muY + 2 \muY \Phi(\muY) < 4 \muY \Phi(\muX),
\)
then $\CV(2) < \CV(1)$ with probability tending to one as $m$ and $n$ increase.
\end{proposition}

\begin{proof}
There are two clusters $G_1$ and $G_2$, where observations from $G_1$ are distributed as
\[	\mathcal{N}\left( \begin{pmatrix} 
    \mu^X \\
    \mu^Y \\
  \end{pmatrix}, \mathbf{I} \right)	\]
and observations from $G_2$ are distributed as
\[	\mathcal{N}\left( \begin{pmatrix} 
    -\mu^X \\
    -\mu^Y \\
  \end{pmatrix}, \mathbf{I} \right)	\]
here $\mu^X$ and $\mu^Y$ are all vectors. Let $G_i$ be the true cluster where observation $i$ is generated from, by assumption
\[	P(G_i=G_1) = P(G_i=G_2) = 1/2	\]

To simplify the notation, let 
\[
 \begin{pmatrix} 
    X_1 \\
    Y_1 \\
  \end{pmatrix}  \sim
  \mathcal{N}\left( \begin{pmatrix} 
    \mu^X \\
    \mu^Y \\
  \end{pmatrix}, \mathbf{I} \right)
\]  
denote the observations from $G_1$ and 
\[
 \begin{pmatrix} 
    X_2 \\
    Y_2 \\
  \end{pmatrix}  \sim
  \mathcal{N}\left( \begin{pmatrix} 
    -\mu^X \\
    -\mu^Y \\
  \end{pmatrix}, \mathbf{I} \right)
\]  
denote the observations from $G_2$

Further, let's denote $\mu^X = \lambda_x e_x$ where $\lambda_x$ is a scalar denotes the distance of $\mu^X$ from origin and $e_x$ is the unit vector point at the same direction as $\mu^X$; $\mu^Y = \lambda_y e_y$ has the same interpretation.

Apply $K$-mean on the $Y$-space, where the two clusters are $\mathcal{N}\left( \mu^Y, \mathbf{I}\right)$ and  $\mathcal{N}\left( -\mu^Y, \mathbf{I}\right)$, the $K$-mean centroids are $\bar{\mu}^Y_1$ and $\bar{\mu}^Y_2$ with $\bar{\mu}^Y_1 = -\bar{\mu}^Y_2$. Note that the boundary between the two clusters are $e^T_y Y > 0 $. So

\begin{align}
\bar{\mu}^Y_1 & = E(Y \mid Y>0) \\
 &= E(Y_1 \mid e^T_y Y_1 > 0) \cdot P(e^T_y Y_1 >0) +E(Y_2 \mid e^T_y Y_2 > 0) \cdot P(e^T_y Y_2 >0)
\end{align}

Note that $e^T_y Y_1$ projects vector $Y_1$ on the direction of $e_y$. And because the $e_y$ is the same direct as $\mu^Y$, it goes through the center of the sphere $\mathcal{N}\left( \mu^Y, \mathbf{I}\right)$. Because the covariance matrix is $\mathbf{I}$, the sphere is symmetric around $e_y$. Therefore, 
\begin{align}
E(Y_1 \mid e^T_y Y_1 = a ) & = a e_y
\end{align}

Also, $Y_1 \sim \mathcal{N}\left( \mu^Y, \mathbf{I}\right)$ so $e^T_y Y_1 \sim \mathcal{N}\left( \lambda_y, 1\right)$. We have

\begin{align}
E(Y_1 \mid e^T_y Y_1 > 0)  & = E \left[  E\left(  Y_1 \mid e^T_y Y_1  \right) \mid e^T_y Y_1 > 0  \right] \\
 from \hspace{0.05in} (3) \hspace{0.05in} above	& =  E( e^T_y Y_1 e_y  \mid e^T_y Y_1 > 0  )  \\ 
	& = e_y E( e^T_y Y_1 \mid e^T_y Y_1 > 0  ) 
\end{align}

Because $e^T_y Y_1 \sim \mathcal{N}\left( \lambda_y, 1\right)  = \lambda_y + Z$, where $Z$ is standard normal, by Lemma $3$ from Appendix C we have 
\begin{align}
E( e^T_y Y_1 \mid e^T_y Y_1 > 0  ) &= E(\lambda_y +  Z \mid Z > -  \lambda_y )  \\
 &= \lambda_y + E(Z \mid Z > -  \lambda_y ) \\
 &= \lambda_y + \frac{\varphi(\lambda_y)}{\Phi(\lambda_y)}
\end{align}

 where $\varphi()$ and $\Phi()$ are the standard normal probability and cumulative distribution function respectively. So, by $(6)$ we have

\begin{align}
E(Y_1 \mid e^T_y Y_1 > 0)  &= \left[ \lambda_y + \frac{\varphi(\lambda_y)}{\Phi(\lambda_y)}  \right] e_y
\end{align}

Similarly, we can have

\begin{align}
E(Y_2 \mid e^T_y Y_2 > 0)  &= - E(Y_1 \mid e^T_y Y_1 < 0) \\
	&= - e_y E( e^T_y Y_1 \mid e^T_y Y_1 < 0  )  \\
	&= \left[ \frac{\varphi(\lambda_y)}{1-\Phi(\lambda_y)} - \lambda_y  \right] e_y 
\end{align}

Because $e^T_y Y_1 \sim \mathcal{N}\left( \lambda_y, 1\right)  = \lambda_y + Z$, it's easy to get 
\begin{align}
P(e^T_y Y_1 > 0) &= P(Z > - \lambda_y) \\
	&= \Phi(\lambda_y) 
\end{align}

By symmetry, we can get 
\begin{align}
P(e^T_y Y_2 > 0) &= 1- \Phi(\lambda_y)
\end{align}

Put everything together, we have

\begin{align}
\bar{\mu}^Y_1 & = E(Y_1 \mid e^T_y Y_1 > 0) \cdot P(e^T_y Y_1 >0) +E(Y_2 \mid e^T_y Y_2 > 0) \cdot P(e^T_y Y_2 >0) \\
 &=  \left[ \lambda_y + \frac{\varphi(\lambda_y)}{\Phi(\lambda_y)}  \right]\cdot \Phi(\lambda_y) e_y + \left[ \frac{\varphi(\lambda_y)}{1-\Phi(\lambda_y)} - \lambda_y  \right]\cdot\left(1- \Phi(\lambda_y)\right) e_y  \\
 & = [2\lambda_y \Phi(\lambda_y) + 2\varphi(\lambda_y) - \lambda_y]e_y 
\end{align}

After training the classifier, because of the identity covariance matrix, the classification boundary is $e^T_x X > 0$. So the $Y$ center for observation with $e^T_x X > 0$ is

\begin{align}
\hat{\mu}^Y_1 &= E(Y_1 \mid e^T_x X_1 > 0)\cdot P(e^T_x X_1 > 0) + E(Y_2 \mid e^T_x X_2 > 0)\cdot P(e^T_x X_2 > 0)\\ 
\text{ X independent of Y}	&= E(Y_1) \cdot P(e^T_x X_1 > 0) + E(Y_2) \cdot P(e^T_x X_2 > 0)\\ 
  & = \mu^Y \cdot P(e^T_x X_1 > 0) -  \mu^Y \cdot P(e^T_x X_2 > 0)\\
  & = \mu^Y \left( P(e^T_x X_1 > 0) -  P(e^T_x X_2 > 0)  \right) \\
   & = \mu^Y \left[ \Phi(\lambda_x) -  (1-\Phi(\lambda_x))  \right] \\
   & = (2\Phi(\lambda_x) - 1)\mu^Y \\
    & = (2\Phi(\lambda_x) - 1)\lambda_y e_y
\end{align}

Because of symmetry and $P(G_i=G_1) = P(G_i=G_2) = 1/2$, it's sufficient to show that for observations with  $e^T_x X > 0$, if $CV(2) < CV(1)$ then the Gabriel CV method correctly picks $k=2$ over $k=1$.

Similar as in the proof of Proposition~\ref{prop:single-general-dim}, by the variance and bias decomposition of MSE, the variance is the same, so only the bias influences the result. Note the predicted center is grand $0$ for $CV(1)$, so to see if  $CV(2) < CV(1)$ one only need to see if $||\bar{\mu}^Y_1-\hat{\mu}^Y_1||^2 < ||\hat{\mu}^Y_1 - 0||^2$, which is true if 
\[ 2\Phi(\lambda_y) + 2 \frac{\varphi(\lambda_y)}{\lambda_y} < 4\Phi(\lambda_x)-1 \] 
this result reduces to the original result of Proposition~\ref{prop:twoclust}
if one sets $\lambda_x = \mu^X$ and $\lambda_y = \mu^Y$.
\end{proof}

\section{Technical Lemmas}
\label{app:technical-lemmas}

\begin{lemma}\label{lem:truncated-normal-moments}

If $Z$ is a standard normal random variable, then
\[
  \E(Z \mid a < Z < b)
    = - \frac{\varphi(b) - \varphi(a)}
             {\Phi(b) - \Phi(a)}
\]
and
\[
  \E\{(Z - \delta)^2 \mid a < Z < b\}
    = \delta^2 + 1
    - \frac{  (b - 2 \delta) \varphi(b)
            - (a - 2 \delta) \varphi(a)}
           {\Phi(b) - \Phi(a)}
\]
for all constants $a$, $b$, and $\delta$, where $\varphi(z)$ and $\Phi(z)$ are
the standard normal probability density and cumulative distribution functions.
These expressions are valid for $a = -\infty$ or $b = \infty$ by taking
limits.

\end{lemma}
\begin{proof}
We will derive the expression for the second moment.  Integrate to get
\begin{align*}
  \E[ (Z - \delta)^2 1\{Z < b\}]
    &= \int_{-\infty}^b (z - \delta)^2 \varphi(z) \, dz \\
    &= (\delta^2 + 1) \Phi(b) - (b - 2 \delta) \varphi(b).
\end{align*}
Now,
\[
  \E\{(Z - \delta)^2 \mid a < Z < b\}
    =
    \frac{  \E[ (Z - \delta)^2 1\{Z < b\}]
          - \E[ (Z - \delta)^2 1\{Z < a\}]}
         { \Phi(b) - \Phi(a) }.
\]
\end{proof}

Lemma~\ref{lem:truncated-normal-moments} has some important special cases:
\begin{align*}
  \E\{Z \mid Z > 0\} &= 2 \varphi(0) = \sqrt{2 / \pi}, \\
  \E\{(Z - \delta)^2 \mid Z > 0 \}
    &= \delta^2 + 1 - 4 \delta \varphi(0), \\
  \E\{(Z - \delta)^2 \mid Z < 0 \}
    &= \delta^2 + 1 + 4 \delta \varphi(0).
\end{align*}

\section{Wold cross-validation}
\label{sec:wold-cv}

In Wold cross-validation, we perform ``speckled'' hold-outs in each fold, leaving out a
random subset of the entries of the data matrix $\dataX \in \R^{N \times P}$. For each value
of $k$ and each fold, we perform the following set of actions to get an
estimate of cross-validation error, $\CV(k)$, which we average over all folds.

\begin{enumerate}
  \item Randomly partition the set of indices $\{ 1, 2, \dotsc, N \} \times \{
    1, 2, \dotsc, P \}$ into a train set $S_\text{train}$ and a test set
    $S_\text{test}$.

  \item Apply a $k$-means fitting procedure that can handle missing data to
    the training data $\{ \dataX_{i,j} : (i,j) \in S_{\text{train}} \}$. This
    gives a set of cluster means $\mu(1), \dotsc, \mu(k) \in \R^P$ and
    cluster labels for the rows, $G_1, G_2, \dotsc, G_N$.

  \item Compute the cross-validation error as
    \[
      \CV(k) = \sum_{(i,j) \in S_{\text{test}}} \{ \dataX_{i,j} - \mu_j(G_i) \}^2,
    \]
    where $\mu_j(G_i)$ denotes the $j$th component of $\mu(G_i)$.
\end{enumerate}

%% \section{Confusion matrix} \label{app:confusion}
%% 
%% \begin{table}
%%   \footnotesize
%% %\captionsetup{justification=centering}
%% \caption{\label{table:confusion-complete} Complete confusion matrix comparing
%% $5$ clusters to $30$ clusters }
%% \begin{tabular}{lccccc|c}
%%   \toprule
%%  & Cluster 1 & Cluster 2 & Cluster 3 & Cluster 4 & Cluster 5 & Total \\ 
%%   \midrule
%%   Cluster 1 & 0 & 0 & 1 & 161 & 2 & 164 \\ 
%%   Cluster 2 & 1 & 0 & 0 & 0 & 185 & 186 \\ 
%%   Cluster 3 & 0 & 0 & 91 & 11 & 2 & 104 \\ 
%%   Cluster 4 & 0 & 102 & 2 & 66 & 0 & 170 \\ 
%%   Cluster 5 & 152 & 0 & 0 & 0 & 0 & 152 \\ 
%%   Cluster 6 & 1 & 2 & 12 & 85 & 4 & 104 \\ 
%%   Cluster 7 & 1 & 10 & 83 & 7 & 0 & 101 \\ 
%%   Cluster 8 & 3 & 145 & 0 & 0 & 0 & 148 \\ 
%%   Cluster 9 & 142 & 0 & 0 & 0 & 5 & 147 \\ 
%%   Cluster 10 & 12 & 27 & 24 & 20 & 6 & 89 \\ 
%%   Cluster 11 & 0 & 5 & 81 & 8 & 0 & 94 \\ 
%%   Cluster 12 & 3 & 68 & 5 & 3 & 1 & 80 \\ 
%%   Cluster 13 & 0 & 1 & 12 & 81 & 5 & 99 \\ 
%%   Cluster 14 & 0 & 1 & 29 & 6 & 38 & 74 \\ 
%%   Cluster 15 & 60 & 53 & 2 & 0 & 0 & 115 \\ 
%%   Cluster 16 & 7 & 0 & 0 & 9 & 83 & 99 \\ 
%%   Cluster 17 & 9 & 0 & 72 & 1 & 1 & 83 \\ 
%%   Cluster 18 & 3 & 16 & 10 & 67 & 5 & 101 \\ 
%%   Cluster 19 & 62 & 2 & 0 & 1 & 8 & 73 \\ 
%%   Cluster 20 & 11 & 0 & 6 & 2 & 65 & 84 \\ 
%%   Cluster 21 & 1 & 13 & 6 & 44 & 6 & 70 \\ 
%%   Cluster 22 & 33 & 15 & 28 & 9 & 0 & 85 \\ 
%%   Cluster 23 & 2 & 18 & 2 & 25 & 22 & 69 \\ 
%%   Cluster 24 & 1 & 38 & 45 & 1 & 0 & 85 \\ 
%%   Cluster 25 & 2 & 57 & 16 & 1 & 0 & 76 \\ 
%%   Cluster 26 & 5 & 3 & 23 & 19 & 0 & 50 \\ 
%%   Cluster 27 & 9 & 0 & 30 & 5 & 20 & 64 \\ 
%%   Cluster 28 & 0 & 7 & 23 & 0 & 38 & 68 \\ 
%%   Cluster 29 & 27 & 2 & 10 & 1 & 11 & 51 \\ 
%%   Cluster 30 & 3 & 5 & 41 & 1 & 10 & 60 \\ \midrule
%%   Total & 550 & 590 & 654 & 634 & 517 & 2945 \\ \bottomrule
%% \end{tabular}
%% \end{table}

\end{document}